---
title: "Architecture: Serialization, Reactive Props, and Reactive Output"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Architecture: Serialization, Reactive Props, and Reactive Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

This vignette explains how `shiny.mantine` actually works under the hood:
how R function calls become a React tree in the browser, how updates flow
back from R to already-mounted components, and the two features that go
beyond simple one-way rendering — a generic reactive-props channel and a
`renderUI()`-equivalent reactive output.

## Why an independent React runtime?

`shiny.mantine` is architecturally inspired by Appsilon's
[shiny.react](https://github.com/Appsilon/shiny.react) — the general idea
of serializing props to JSON, syncing input values via
`Shiny.setInputValue()`, and loading assets through an `htmlDependency` is
the same. But it does **not** reuse `shiny.react`'s compiled JavaScript
runtime or its React copy.

The reason is a version conflict: Mantine v9 requires React >= 19.2, while
`shiny.react` shares a single React 18.3.1 copy across every package that
uses its `externals` mechanism. Mounting two different React copies in
the same page causes hook-dispatcher conflicts ("Invalid hook call") —
see [Appsilon/shiny.react#87](https://github.com/Appsilon/shiny.react/issues/87).
So `shiny.mantine` bundles its own React 19 / ReactDOM 19 copy inside
`inst/www/mantine.js`, completely independent from `shiny.react`'s
runtime, with its own small mount and input-sync layer reimplementing the
same concepts.

## From R call to React element

Every component wrapper is a thin call to one of two internal helpers in
`R/mantine-element.R`:

```{r}
# displayComponent() is used for components with no special behavior:
# it just forwards whatever arguments it receives.
Card
```

```{r}
# Some components validate/transform an argument before forwarding it —
# e.g. Button() requires both a label and an inputId.
Button
```

Both ultimately call `mantineElement()`, which builds a plain, serializable
tree — `list(type = "element", name = <string>, props = <list>)` — tagged
with the S3 class `mantine_element`:

```{r}
btn <- Button("Click me", inputId = "btn1", color = "blue")
str(unclass(btn))
```

Each prop value is recursively normalized by `toMantineData()` into one
of a small set of tagged shapes (`raw`, `array`, `object`, `element`,
`html`) that the JS side's `buildElement()`/`buildProps()`
(`js/src/serialization.js`) knows how to turn back into real React
elements or plain JS values:

```{r}
card <- Card(withBorder = TRUE, Text("Hello"))
str(unclass(card), max.level = 3)
```

Nesting works exactly like `htmltools`: any unnamed argument becomes part
of the element's children, and a `mantine_element` passed as an argument
is itself serialized recursively (that's how `Text("Hello")` ends up
nested inside `Card()`'s `children` prop above).

`MantineProvider()` is the one function that doesn't return a plain
`mantine_element` — it calls `renderMantineRoot()`, which serializes the
tree to JSON and wraps it in a self-mounting `htmltools` fragment (a
`<div>` holding the JSON plus a `<script>` that calls
`window.jsmodule['@/shiny.mantine'].mount()`). This is why
`MantineProvider()` must always be the outermost call: it's the one that
actually produces mountable HTML output.

## The generic reactive-props channel

Dedicated `update*()` functions like `updateMantineTextInput()` only ever
touch one thing: an input's `value`. But plenty of props aren't a
"value" in that sense — a `Modal`'s `opened`, a `Stepper`'s `active`, a
`Badge`'s `color`. For these, there's a single generic mechanism:
`updateMantineProps()`.

```{r}
updateMantineProps
```

It sends `list(id = <mantineId>, props = list(...))` over a
`shinyMantineUpdateProps` custom message. On the JS side,
`withReactiveProps()` — a small HOC wrapping every component that accepts
a `mantineId` prop — keeps a local "patch" object in React state per
mounted `mantineId`, and merges any incoming props into it, triggering a
re-render:

```js
function withReactiveProps(Component) {
  return function Wrapped({ mantineId, ...props }) {
    const [patch, setPatch] = useState({});
    useEffect(() => {
      if (!mantineId) return undefined;
      propUpdateHandlers[mantineId] = (newProps) => setPatch((prev) => ({ ...prev, ...newProps }));
      return () => { delete propUpdateHandlers[mantineId]; };
    }, [mantineId]);
    return React.createElement(Component, { ...props, ...patch });
  };
}
```

This is what makes `Modal(mantineId = "my_modal", ...)` +
`updateMantineProps(session, "my_modal", opened = TRUE)` work, and it's
reused throughout the package: `Collapse`'s `opened`, `Stepper`'s
`active`, or any component's `disabled`/`color`/`label` can all be changed
this way without rebuilding the tree.

```r
# ui:
Modal(mantineId = "confirm", inputId = "confirm_state", title = "Confirm", Text("Are you sure?"))

# server:
observeEvent(input$open_btn, {
  updateMantineProps(session, "confirm", opened = TRUE)
})
```

Not every component supports `mantineId` — only ones explicitly wrapped
with `withReactiveProps()` in `js/src/index.js`, which in practice is
almost everything in the registry (including every component the
automated generator produces, see below).

## Reactive output: `mantineOutput()` / `renderMantine()`

`MantineProvider()` mounts its whole tree once, statically, at page load.
There is no built-in way to make it reactively rebuild — unlike
`uiOutput()`/`renderUI()` for regular HTML. `mantineOutput()`/`renderMantine()`
close this gap:

```r
# ui:
mantineOutput("summary")

# server:
output$summary <- renderMantine({
  Card(withBorder = TRUE, Text(paste("Total:", sum(input$values))))
})
```

Each `mantineOutput()` mounts its own independent React root via a
generic Shiny output binding registered in `js/src/index.js`. As with
`renderUI()`, every re-evaluation of the expression rebuilds the whole
sub-tree, so uncontrolled local state inside it (e.g. text typed into a
`TextInput` with no stable `inputId`) can be lost between updates.

If you want Mantine's theme/color-scheme inside the reactive output too,
you can wrap the content in a nested `MantineProvider()` — this is
normally impossible to nest usefully (it would just serialize as an inert
HTML string, since its embedded mount `<script>` would never execute once
injected via `dangerouslySetInnerHTML`), but `toMantineData()` special-cases
it: `renderMantineRoot()` attaches the pre-serialization element as an R
attribute (`mantine_root`), and `toMantineData()` checks for and recovers
it before falling back to generic HTML serialization:

```{r}
# renderMantineRoot()/toMantineData() are internal (not exported), since
# ordinary usage never needs to call them directly.
nested <- shiny.mantine:::renderMantineRoot(Stack(Text("nested")))
recovered <- shiny.mantine:::toMantineData(nested)
str(recovered, max.level = 1)
```

## The automated component generator

Most Mantine components need no special Shiny wiring at all — they're
"pure display": props and children in, rendered output out, with no
value to sync. Writing a hand-rolled adapter for each of these would be
pure repetition, so `js/scripts/generate-components.js` generates them
from a declarative manifest instead.

Each manifest entry looks like this:

```js
{ name: 'Blockquote', slug: 'blockquote', doc: 'Styled quotation block' },
{
  name: 'List',
  slug: 'list',
  doc: 'Bulleted/numbered list',
  compound: [{ export: 'ListItem', jsName: 'Item' }],
},
```

Running the script regenerates two matched files: `js/src/generated-components.js`
(imports the named exports from `@mantine/core` and registers them, each
wrapped in `withReactiveProps()`) and `R/generated-components.R` (an R
`displayComponent()` wrapper with a roxygen block, one per manifest
entry). To add a new "display only" component to the package:

```
cd js
# edit the MANIFEST array in scripts/generate-components.js
node scripts/generate-components.js
npm run build
```

```r
devtools::document()
devtools::install()
```

Components that need real interaction — a stateful value, a click that
should report to Shiny, an anchored dropdown, a controller object like
`Tree`'s `useTree()` — are still written by hand (see `R/Inputs2.R`,
`R/Overlays.R`, `R/AdvancedComponents.R`, and the corresponding adapters
in `js/src/index.js`), since that's exactly the part a declarative
manifest can't express.
