---
title: "Component Gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Component Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

A categorized tour of every component family in the package. Each section
is a runnable snippet you can drop into a `shinyApp(ui, server)` call.
For the eight satellite packages (dates, notifications, modals,
spotlight, charts, code-highlight, nprogress, tiptap), see
`vignette("satellite-packages")` instead — they're not repeated here.

## Layout

`Group()` (horizontal flex), `Stack()` (vertical flex), `SimpleGrid()`
(uniform-column responsive grid), `Grid()`/`GridCol()` (12-column grid),
`Container()`, `Box()`, `Center()`, `Splitter()`/`SplitterPanel()`
(resizable panes), `AspectRatio`, `Flex`, `Space`:

```r
Stack(
  Group(Text("Left"), Text("Right")),
  SimpleGrid(cols = 3, Card("A"), Card("B"), Card("C")),
  Grid(
    GridCol(span = 8, Text("Main content")),
    GridCol(span = 4, Text("Sidebar"))
  ),
  Splitter(
    style = list(height = "200px"),
    SplitterPanel(Text("Left pane")),
    SplitterPanel(Text("Right pane"))
  )
)
```

## Typography & feedback

`Text()`, `Title()`, `Blockquote`, `Highlight`, `Code`, `Mark`,
`Badge()`, `Alert`, `Notification` (see the notifications satellite
package), `Divider()`, `Indicator`, `Loader`, `Skeleton`, `Kbd`,
`Spoiler`, `Typography` (applies prose styling to arbitrary HTML),
`VisuallyHidden`:

```r
Stack(
  Title("Page title", order = 2),
  Text("Regular text", c = "dimmed"),
  Badge("New", color = "green"),
  Alert("Something needs your attention", title = "Heads up", color = "yellow"),
  Highlight(highlight = "Mantine", "Built with Mantine UI"),
  Spoiler(maxHeight = 40, showLabel = "Show more", hideLabel = "Hide", "A long paragraph...")
)
```

## Navigation

`NavLink()`/`navLinkItem()`, `Burger()`/`navbarBurger()`, `Tabs()` family
(page-aware, see `vignette("getting-started")`), `Menu()` family
(including submenus, checkbox/radio items, search, and a right-click
context-menu variant), `Menubar()` (a desktop-app-style menu bar,
distinct from a single dropdown `Menu()`), `Breadcrumbs`, `Pagination()`,
`Stepper()` (multi-step wizard), `Tree()` (hierarchical data):

```r
Stack(
  Menu(
    MenuTarget(Button("Actions", inputId = "menu_btn")),
    MenuDropdown(
      menuItem("action", "edit", "Edit"),
      MenuDivider(),
      menuItem("action", "delete", "Delete")
    )
  ),
  Stepper(
    mantineId = "wizard", inputId = "wizard_click", active = 0,
    StepperStep(label = "Account", description = "Create your account"),
    StepperStep(label = "Details", description = "Add your details"),
    StepperCompleted("All done!")
  ),
  Tree(
    inputId = "tree_click",
    data = list(
      list(value = "src", label = "src", children = list(
        list(value = "src/app.R", label = "app.R")
      )),
      list(value = "readme", label = "README.md")
    )
  )
)
```

`Stepper()`'s `active` step is controlled entirely by the app — advance
it with `updateMantineProps(session, "wizard", active = <n>)` in response
to whatever validation/logic decides the step is complete.

## Inputs

Text-like: `TextInput()`, `Textarea()`, `PasswordInput()`, `Autocomplete()`,
`MaskInput()` (masked text, e.g. phone numbers). Numeric: `NumberInput()`,
`Slider()`/`RangeSlider()`, `Rating()`. Choice: `Select()`, `MultiSelect()`,
`TagsInput()`, `NativeSelect()`, `TreeSelect()` (hierarchical data),
`SegmentedControl()`. Boolean/group: `Switch()`, `Checkbox()`/`CheckboxGroup()`,
`Radio()`/`RadioGroup()`, `Chip()`/`ChipGroup()`, `CheckboxCard()`/`RadioCard()`
(card-styled selectable items). Other: `PinInput()` (OTP-style),
`JsonInput()`, `ColorInput()`/`ColorPicker()`, `FileInput()`/`FileButton()`
(metadata-only file selection), `Accordion()` family.

```r
Stack(
  TextInput(inputId = "name", label = "Name"),
  MaskInput(inputId = "phone", mask = "+1 (999) 999-9999", label = "Phone"),
  Select(inputId = "country", label = "Country", data = c("US", "IT", "FR")),
  CheckboxGroup(
    inputId = "fruits", label = "Pick fruits",
    CheckboxGroupItem(value = "apple", label = "Apple"),
    CheckboxGroupItem(value = "banana", label = "Banana")
  ),
  RadioGroup(
    inputId = "plan", label = "Plan",
    RadioCard(value = "free", Text("Free")),
    RadioCard(value = "pro", Text("Pro"))
  )
)
```

Note the naming split for group inputs: the group itself
(`CheckboxGroup()`/`SwitchGroup()`) is stateful and reports an array of
selected values, but its children must be the dedicated
`CheckboxGroupItem()`/`SwitchGroupItem()` (or, for radio/chip groups, the
plain `Radio()`/`Chip()`) — never the standalone `Checkbox()`/`Switch()`,
which manage their own independent checked state and would conflict with
the group's.

## Overlays

`Modal()`, `Drawer()`, `Dialog()`, `Popover()` (+`PopoverTarget()`/`PopoverDropdown()`),
`Affix()`, `LoadingOverlay()`, `Collapse()` — all controlled from R via
`mantineId` + `updateMantineProps()`:

```r
Stack(
  Button("Open", inputId = "open_btn"),
  Modal(
    mantineId = "demo_modal", inputId = "modal_state", title = "Demo",
    Text("Modal content")
  ),
  Collapse(mantineId = "details", opened = FALSE, Text("Collapsible content"))
)

# server:
observeEvent(input$open_btn, {
  updateMantineProps(session, "demo_modal", opened = TRUE)
  updateMantineProps(session, "details", opened = TRUE)
})
```

More than one `Modal()`/`Drawer()` open at the same time needs
`ModalStack()`/`DrawerStack()` instead of standalone ones — each manages
its own `opened` boolean independently and doesn't coordinate z-index or
focus trapping with the others, so two open together would render with
mismatched layering. Wrapping the same `Modal()`s in `ModalStack()`
delegates that coordination to Mantine; every child keeps opening and
closing exactly as above, and the whole stack can be closed at once with
`updateMantineProps(session, stackId, closeAll = TRUE)`:

```r
ModalStack(
  mantineId = "delete_stack",
  Modal("delete_page", inputId = "delete_page_state", title = "Delete this page?",
    Button("Continue", inputId = "go_to_confirm")
  ),
  Modal("confirm_delete", title = "Are you really sure?",
    Button("Yes, delete", inputId = "confirm_delete_btn", color = "red")
  )
)

# server:
observeEvent(input$go_to_confirm, {
  updateMantineProps(session, "confirm_delete", opened = TRUE)
})
observeEvent(input$confirm_delete_btn, {
  updateMantineProps(session, "delete_stack", closeAll = TRUE)
})
```

For full control over a modal/drawer's internal layout (e.g. extra
controls next to the title), use `ModalRoot()`/`DrawerRoot()` — the
lower-level piece behind `Modal()`/`Drawer()` — composed with
`ModalOverlay()`/`ModalContent()`/`ModalHeader()`/`ModalTitle()`/
`ModalCloseButton()`/`ModalBody()` (and the `Drawer*` equivalents). They
open/close from R exactly like the all-in-one versions, including
`ModalStack()`/`DrawerStack()` participation:

```r
ModalRoot(
  mantineId = "custom_modal", inputId = "custom_modal_state",
  ModalOverlay(),
  ModalContent(
    ModalHeader(ModalTitle("Custom layout"), ModalCloseButton()),
    ModalBody(Text("Modal content"))
  )
)
```

`FocusTrap()`/`RemoveScroll()` are the same focus-trapping/scroll-locking
mechanisms `Modal()`/`Drawer()` use internally, for building other fully
custom overlays that need the same behavior.

## Data display

`Table()` family (compose by hand for full control), `DataTable()`
(search/sort/selection built in), `DataList` (key/value pairs),
`EmptyState` (placeholder for empty views), `Avatar()`/`AvatarGroup`,
`Timeline`, `List`, `Card()`/`CardSection()`, `Image()`, `RingProgress()`,
`SemiCircleProgress`, `NumberFormatter`, `Pill`/`PillGroup`, `Marquee`,
`RollingNumber`, `ColorSwatch`, `Overlay`:

```r
DataTable(
  "team",
  data = list(
    list(value = 1, name = "Ada", role = "Engineer"),
    list(value = 2, name = "Grace", role = "Researcher")
  ),
  columns = list(list(key = "name", label = "Name"), list(key = "role", label = "Role")),
  searchable = TRUE, sortable = TRUE, selectable = TRUE
)
```

## Drag & drop, dropzone, carousel

`SortableList()`/`SortableTable()` (reorder via drag & drop, report the
new order), `Dropzone()` (drag & drop upload, metadata only),
`FileButton()` (a button that opens the native file picker, metadata
only), `Carousel()`/`CarouselSlide()`:

```r
Stack(
  SortableList(
    "elements", withHandle = TRUE,
    items = list(
      list(value = "H", label = "Hydrogen"),
      list(value = "He", label = "Helium")
    )
  ),
  Dropzone(inputId = "upload", accept = "image/png,image/jpeg"),
  FileButton(inputId = "avatar", label = "Upload avatar", accept = "image/*")
)
```

## Provider components

`MantineProvider()` (required, always outermost), `DatesProvider()`
(locale/date format), `DirectionProvider()` (`"ltr"`/`"rtl"` reading
direction):

```r
MantineProvider(
  DirectionProvider(dir = "ltr"),
  DatesProvider(settings = list(locale = "en")),
  # ... your app ...
)
```

`MantineThemeProvider(theme = list(...))` updates the theme *object*
exposed via Mantine's `useMantineTheme()` hook for everything nested
inside. **It does not restyle standard components** (`Button()`,
`Badge()`, ...): those get their colors from CSS variables that only the
outermost `MantineProvider()` injects once, so nesting a
`MantineThemeProvider()` with a different `primaryColor` will *not* turn
a `Button()` a different color — that component was already compiled
against the outer theme's CSS variables. Its practical use in
`shiny.mantine` is narrow: it exists mainly for API completeness and for
any future/custom component that reads `theme` reactively through that
hook, not as a scoped-restyling tool. If you need a visually different
theme for part of a page, that currently requires a second, independent
`MantineProvider()` (with its own color-scheme-sync caveat — see the
README's "Known limitations").

## Additional components

Three coverage passes over every export of `@mantine/core` (plus the
satellite packages) closed most remaining gaps — mainly in
`R/AdvancedComponents.R`, `R/Charts.R`, `R/Dates.R`, and
`R/CodeHighlight.R`. Each was verified in-browser (Playwright, zero
console errors):

- **Navigation/structure**: `Stepper()` (multi-step wizard, `active`
  controlled via `updateMantineProps()`), `Tree()`/`TreeSelect()`
  (hierarchical data, wired to Mantine's `useTree()` internally),
  `Collapse()` (animated show/hide, toggled via `updateMantineProps()`),
  `Menubar()`, `Menu()` extensions (submenus, checkbox/radio items,
  search, context-menu trigger), `Splitter()`/`SplitterPanel()`
  (resizable panes), `CardSection()`/`ActionIconGroup()`. For a fully
  custom arrangement of `Pagination()`'s prev/next/first/last controls,
  `PaginationRoot()` + `PaginationFirst()`/`PaginationPrevious()`/
  `PaginationItems()`/`PaginationNext()`/`PaginationLast()`/
  `PaginationDots()` read the current page from it automatically (no
  per-part wiring needed) and stay stateful exactly like `Pagination()`.
- **Inputs**: `CheckboxGroup()`/`SwitchGroup()` (array-valued group
  inputs, same pattern as `RadioGroup()`/`ChipGroup()`),
  `CheckboxCard()`/`RadioCard()`, `FileButton()`, `MaskInput()` (masked
  text input, mounted uncontrolled internally since Mantine's masking
  hook mutates the DOM value directly), `DirectionProvider()`.
  `HueSlider()`/`AlphaSlider()`/`AngleSlider()` are standalone versions
  of the channel sliders `ColorPicker()` uses internally (plus a
  circular 0-360° dial for `AngleSlider()`). `RadioIndicator()`/
  `CheckboxIndicator()` are the purely decorative "checked" look
  `Radio()`/`Checkbox()` render, usable without a real input behind them
  (e.g. a read-only status list). `Input()`/`InputBase()` provide
  Mantine's styled box (border, focus ring, left/right sections) with
  *no* value/onChange management of their own — unlike `TextInput()` and
  friends, not synced to Shiny; use them for visual composition.
  `InputWrapper()`/`InputLabel()`/`InputDescription()`/`InputError()`/
  `InputPlaceholder()` give arbitrary custom content the same
  label/description/error chrome.
- **Charts** (`@mantine/charts`, beyond `LineChart()`/`BarChart()`/
  `AreaChart()`/`PieChart()`/`DonutChart()`): `RadarChart()`,
  `CompositeChart()`, `RadialBarChart()`, `BubbleChart()`,
  `FunnelChart()`, `Sparkline()`, `ScatterChart()`, `Treemap()`,
  `Heatmap()`, `SankeyChart()` — see `vignette("satellite-packages")` for
  each one's `data` shape. Charts in an unbounded container (no fixed
  height) need an explicit `h`; several of these newer chart types don't
  ship a default the way the original five do.
- **Dates** (`@mantine/dates`, beyond the pickers in
  `vignette("satellite-packages")`): `TimePicker()`, `TimeGrid()`,
  `MiniCalendar()`, `InlineDateTimePicker()`.
- **`CodeHighlightTabs()`** (`@mantine/code-highlight`) — a tabbed,
  multi-file code block, same plain-text adapter as `CodeHighlight()`.
- **Display**: `Notification()` (a static notification box for your own
  layout — distinct from the toast popups in `vignette("satellite-packages")`),
  `Transition()` (mount/unmount animation — Mantine's version takes a
  render-prop `children`, incompatible with R's data-only `...`, so it's
  wrapped internally to accept plain children instead), `Portal()`,
  `ScrollAreaAutosize()`/`NativeScrollArea()`, and the multi-segment
  `ProgressRoot()`/`ProgressSection()`/`ProgressLabel()` compound form of
  `Progress()`.
- **Generated passthrough components** (see `vignette("architecture")`'s
  section on the automated generator): `Fieldset`, `DataList` family,
  `EmptyState` family, `Marquee`, `NumberFormatter`, `Pill`/`PillGroup`,
  `RollingNumber`, `SemiCircleProgress`, `TooltipFloating`, `TooltipGroup`,
  `Typography`, `VisuallyHidden`, the `Modal`/`Drawer`/`Pagination`
  compound sub-parts and `Input` family above, plus small standalone
  pieces Mantine renders internally: `CheckIcon()`, `CloseIcon()`,
  `AccordionChevron()`, `RadioIcon()`, `ActionIconGroupSection()`,
  `ButtonGroupSection()`.

### Intentionally out of scope

A previous version of this section excluded the `Modal`/`Drawer`/
`Pagination`/`Spotlight` compound forms, `HueSlider`/`AlphaSlider`/
`AngleSlider`, `FocusTrap`/`RemoveScroll`, `MantineThemeProvider`, the
`Input` family, and a handful of small internal icons/sections
(`CheckIcon`, `CloseIcon`, `RadioIcon`, `RadioIndicator`,
`CheckboxIndicator`, `AccordionChevron`, `ActionIconGroupSection`,
`ButtonGroupSection`). On closer inspection none of these actually needed
function-valued props or non-serializable values (refs, live positioning
context) — they're now all wrapped, and documented above and in
`vignette("satellite-packages")`.

What remains genuinely out of scope has a real architectural reason, not
just "not gotten to yet":

- **`Combobox` internals** (`ComboboxDropdown`, `ComboboxOption`, ...) —
  driven by Mantine's own `useCombobox()` hook (open state, active
  option, keyboard navigation state), which can't be expressed through
  R's data-only props; `Select()`/`MultiSelect()`/`Autocomplete()`/
  `TagsInput()`/every stateful input already built on top of it are
  exposed at the higher level.
- **`FloatingArrow`/`FloatingIndicator`** — need a live floating-UI
  `context` object and/or a target-element `ref`, neither of which is a
  plain, JSON-serializable R value.
- **SSR/Next.js-specific helpers**: `ColorSchemeScript`,
  `HeadlessMantineProvider`, `InlineStyles` — genuinely not applicable to
  Shiny's client-side-only mount architecture (there's no server-rendered
  HTML to hydrate, and no reason to run Mantine without emitting its own
  CSS).
- **`TableOfContents`/`OverflowList`** — require scanning real page DOM
  headings or render-prop-heavy overflow logic, both a poor fit for a
  declarative R wrapper.
- **`Calendar`** and its sub-parts (`Month`, `Day`, `WeekdaysRow`,
  `YearLevel`, `DecadeLevel`, ...) — the low-level primitive
  `DatePicker()`/`DatePickerInput()` are built on; customization goes
  through function-based `getDayProps`, a poor fit for R's data-only props.

## Where to go next

- `vignette("architecture")` — how these components are serialized and
  kept in sync with Shiny.
- `vignette("satellite-packages")` — dates, notifications, modals,
  spotlight, charts, code-highlight, nprogress, and the rich text editor.
