---
title: "Core: Typography"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Typography}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Typography" category](https://mantine.dev/core/text/)
on mantine.dev/core. `Table()` is covered in `vignette("core-data-display")`
instead, matching mantine.dev's own categorization of it as data display.

## Text

<https://mantine.dev/core/text/> — the base text component; every
size/color/weight prop (`size`, `c` for color, `fw`, `ta`, `truncate`,
...) works exactly as documented on mantine.dev.

```r
Text("Regular paragraph text.")
Text("Dimmed, smaller helper text.", size = "sm", c = "dimmed")
Text("Bold, colored text.", fw = 700, c = "blue")
```

`Text()` also supports a gradient variant:

```r
Text("Gradient heading", variant = "gradient", gradient = list(from = "blue", to = "cyan"), fw = 900, size = "xl")
```

## Title

<https://mantine.dev/core/title/> — semantic headings (`order` 1-6,
mapping to `<h1>`-`<h6>`).

```r
Stack(Title("Heading 1", order = 1), Title("Heading 2", order = 2), Title("Heading 3", order = 3))
```

## Blockquote

<https://mantine.dev/core/blockquote/> — a styled quotation block, with
an optional `cite` and `icon`.

```r
Blockquote(cite = "– Ada Lovelace", icon = IconFingerprint(size = 20), "The Analytical Engine has no pretensions whatever to originate anything.")
```

## Code

<https://mantine.dev/core/code/> — inline or block code-styled text
(for real syntax-highlighted code blocks, see `CodeHighlight()` in
`vignette("satellite-packages")` instead).

```r
Text("Run ", Code("install.packages(\"shiny\")"), " to get started.")
Code(block = TRUE, "shiny::runApp(\"app.R\")")
```

## Highlight

<https://mantine.dev/core/highlight/> — renders text with one or more
substrings visually highlighted.

```r
Highlight(highlight = c("Mantine", "Shiny"), "shiny.mantine brings Mantine UI to Shiny apps.")
```

## List

<https://mantine.dev/core/list/> — a bulleted or numbered list;
`ListItem()`s nest inside.

```r
List(type = "ordered", ListItem("First step"), ListItem("Second step"), ListItem("Third step"))
```

## Mark

<https://mantine.dev/core/mark/> — highlighted text using the native
`<mark>` element (a simpler, single-purpose alternative to `Highlight()`
when you already know exactly which text node to mark).

```r
Text("This is ", Mark("marked"), " text.")
```

## Typography

<https://mantine.dev/core/typography/> — applies Mantine's prose
styling (headings, links, lists, code blocks, ...) to arbitrary raw HTML
content, e.g. Markdown rendered server-side.

```r
Typography(HTML(markdown::markdownToHTML(text = "# Title\n\nSome **bold** text and a [link](https://mantine.dev).", fragment.only = TRUE)))
```

## Where to go next

- `vignette("core-data-display")` — `Table()` and other components that
  display structured data.
- `vignette("core-misc")` — `Divider()`, `Paper()`, and other
  miscellaneous layout/display primitives.
