---
title: "Core: Combobox"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Combobox}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Combobox" category](https://mantine.dev/core/select/)
on mantine.dev/core — every dropdown-driven choice input, all built on
top of Mantine's internal `useCombobox()` state machine (open state,
active option, keyboard navigation). That hook itself can't be expressed
through R's plain serializable props directly — but unlike earlier
versions of this package, it no longer has to be: `Combobox()` (below)
creates and owns the store on the JS side for you, so even the headless
primitive stays available from R for building a fully custom
dropdown-driven input, on top of the ready-made ones below.

## Select

<https://mantine.dev/core/select/> — a single-choice searchable dropdown.
`data` is a plain character vector, or a list of `list(value=, label=)`
for a different label than the stored value.

```r
Select(
  inputId = "country", label = "Country",
  data = c("Italy", "France", "Germany", "Spain"),
  searchable = TRUE, placeholder = "Pick a country"
)
```

## MultiSelect

<https://mantine.dev/core/multi-select/> — same idea, multiple
selections; `input[[inputId]]` is a character vector.

```r
MultiSelect(
  inputId = "skills", label = "Skills",
  data = c("R", "Python", "JavaScript", "SQL"),
  value = list("R")
)
```

## TagsInput

<https://mantine.dev/core/tags-input/> — like `MultiSelect()`, but the
user can also type and add values not present in `data`.

```r
TagsInput(inputId = "tags", label = "Tags", data = c("bug", "feature", "docs"))
```

## Autocomplete

<https://mantine.dev/core/autocomplete/> — a free-text field with
suggestions; unlike `Select()`, the user isn't restricted to `data`'s
values.

```r
Autocomplete(inputId = "city", label = "City", data = c("Rome", "Milan", "Naples", "Turin"))
```

## TreeSelect

<https://mantine.dev/core/tree-select/> (a `shiny.mantine`-specific
combination of `Select()`'s dropdown with `Tree()`'s hierarchical `data`
— not a standalone mantine.dev page, but built the same way as the
others here) — a dropdown for picking one node out of nested data:

```r
TreeSelect(
  inputId = "folder", label = "Destination folder",
  data = list(
    list(value = "docs", label = "Documents", children = list(
      list(value = "docs/reports", label = "Reports")
    )),
    list(value = "images", label = "Images")
  )
)
```

## Pill

<https://mantine.dev/core/pill/> — a small rounded tag element, the
visual building block `MultiSelect()`/`TagsInput()` render for each
selected value; `PillGroup()` wraps several together. Useful standalone
for displaying (not necessarily editing) a list of tags:

```r
PillGroup(Pill("R"), Pill("Shiny"), Pill("Mantine", withRemoveButton = TRUE))
```

## ComboboxPopover

<https://mantine.dev/core/combobox-popover/> — like `Select()`, but
renders no input of its own: you supply the clickable target (typically
a `Button()`) via `ComboboxPopoverTarget()`. Useful when the trigger
shouldn't look like a text field.

```r
ComboboxPopover(
  inputId = "framework", data = c("React", "Angular", "Vue", "Svelte"),
  ComboboxPopoverTarget(Button("Pick a framework", variant = "default"))
)
```

## Combobox (headless primitive)

<https://mantine.dev/core/combobox/> — reach for this only when
`Select()`/`MultiSelect()`/`Autocomplete()`/`TagsInput()`/
`ComboboxPopover()` don't fit (e.g. options need fully custom
icons/layout beyond what `renderOption`-style composition in those
covers). The real Mantine `Combobox` is driven by a live `useCombobox()`
store (a JS object with methods); `Combobox()` creates and owns that
store internally, so the R API stays declarative:

- `ComboboxTarget()`'s single child (e.g. a `Button()`) toggles the
  dropdown on click automatically.
- Selecting a `ComboboxOption()` sends its `value` to `input[[inputId]]`
  (as an event, like `actionButton()`) and closes the dropdown;
  highlighting the current selection (`active = TRUE`) is your
  responsibility, same as Mantine's own controlled examples.
- `ComboboxSearch()` is wired like any text input
  (`input[[inputId]]` on every keystroke) — pair it with
  `renderMantine()`/`mantineOutput()` to re-render `ComboboxOptions()`
  server-side as the user types, since an R-side `filter` function can't
  cross the bridge the way Mantine's own `filter` prop does.

```r
Combobox(
  inputId = "fruit",
  ComboboxTarget(Button("Pick a fruit", variant = "default")),
  ComboboxDropdown(
    ComboboxOptions(
      ComboboxOption(value = "apple", "🍎 Apple"),
      ComboboxOption(value = "banana", "🍌 Banana")
    )
  )
)

# server:
observeEvent(input$fruit, {
  message("Picked: ", input$fruit)
})
```

## Where to go next

- `vignette("core-inputs")` — text/numeric/choice/boolean form inputs.
- `vignette("core-misc")` — the full "Intentionally out of scope" list,
  covering the parts of `Combobox`'s design (a raw `ref`/DOM-driven
  `FloatingIndicator`, `Calendar`'s `getDayProps`) that stay out of
  reach even with this pragmatic wrapper.
