---
title: "Core: Combobox"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Combobox}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Combobox" category](https://mantine.dev/core/select/)
on mantine.dev/core — every dropdown-driven choice input, all built on
top of Mantine's internal `useCombobox()` state machine (open state,
active option, keyboard navigation). That hook itself isn't exposed from
R (its state can't be expressed as plain serializable props — see
`vignette("core-misc")`'s "Intentionally out of scope" section), but
every component built on top of it is fully wrapped and stateful.

## Select

<https://mantine.dev/core/select/> — a single-choice searchable dropdown.
`data` is a plain character vector, or a list of `list(value=, label=)`
for a different label than the stored value.

```r
Select(
  inputId = "country", label = "Country",
  data = c("Italy", "France", "Germany", "Spain"),
  searchable = TRUE, placeholder = "Pick a country"
)
```

## MultiSelect

<https://mantine.dev/core/multi-select/> — same idea, multiple
selections; `input[[inputId]]` is a character vector.

```r
MultiSelect(
  inputId = "skills", label = "Skills",
  data = c("R", "Python", "JavaScript", "SQL"),
  value = list("R")
)
```

## TagsInput

<https://mantine.dev/core/tags-input/> — like `MultiSelect()`, but the
user can also type and add values not present in `data`.

```r
TagsInput(inputId = "tags", label = "Tags", data = c("bug", "feature", "docs"))
```

## Autocomplete

<https://mantine.dev/core/autocomplete/> — a free-text field with
suggestions; unlike `Select()`, the user isn't restricted to `data`'s
values.

```r
Autocomplete(inputId = "city", label = "City", data = c("Rome", "Milan", "Naples", "Turin"))
```

## TreeSelect

<https://mantine.dev/core/tree-select/> (a `shiny.mantine`-specific
combination of `Select()`'s dropdown with `Tree()`'s hierarchical `data`
— not a standalone mantine.dev page, but built the same way as the
others here) — a dropdown for picking one node out of nested data:

```r
TreeSelect(
  inputId = "folder", label = "Destination folder",
  data = list(
    list(value = "docs", label = "Documents", children = list(
      list(value = "docs/reports", label = "Reports")
    )),
    list(value = "images", label = "Images")
  )
)
```

## Pill

<https://mantine.dev/core/pill/> — a small rounded tag element, the
visual building block `MultiSelect()`/`TagsInput()` render for each
selected value; `PillGroup()` wraps several together. Useful standalone
for displaying (not necessarily editing) a list of tags:

```r
PillGroup(Pill("R"), Pill("Shiny"), Pill("Mantine", withRemoveButton = TRUE))
```

## Combobox internals (not wrapped)

`Combobox`, `ComboboxDropdown`, `ComboboxOption`, `ComboboxSearch`, and
the rest of its compound sub-parts are intentionally not wrapped: they
require driving `useCombobox()` yourself (tracking open state, the
active/highlighted option, keyboard events), which has no
plain-serializable-props equivalent. Every component on this page is
already built on top of it — reach for those first, and only consider
this a hard limit if you need a completely bespoke dropdown-driven input
that none of them cover.

## Where to go next

- `vignette("core-inputs")` — text/numeric/choice/boolean form inputs.
- `vignette("core-misc")` — the full "Intentionally out of scope" list,
  including why `Combobox` internals aren't wrapped.
