---
title: "Core: Buttons"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Buttons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Buttons" category](https://mantine.dev/core/button/)
on mantine.dev/core. For the *recipes* shown on
[ui.mantine.dev/category/buttons](https://ui.mantine.dev/category/buttons/)
(color scheme toggle, copy-to-clipboard, split button, ...), see
`ButtonWithMenu()`/`SplitButton()`/`SocialButton()`/`ColorSchemeToggle()`/
`LoadingProgressButton()` in `vignette("extras")` and the runnable
`inst/examples/buttons-app.R` demo instead — those are compositions on
top of the plain components below, not separate mantine.dev pages.

## Button

<https://mantine.dev/core/button/> — the standard button.
`inputId` makes it stateful like `shiny::actionButton()`:
`input[[inputId]]` is a click counter, incremented on every click (there
is no synced "value" otherwise, so there's no `updateButton()`).

```r
Button("Click me", inputId = "my_btn", variant = "filled", color = "blue", leftSection = IconPlus(size = 16))

# server:
observeEvent(input$my_btn, {
  message("Clicked ", input$my_btn, " times")
})
```

`Button.Group` (`ButtonGroup()`) visually joins several buttons together
with shared borders; `ButtonGroupSection()` adds a non-interactive
section (e.g. a label) between them:

```r
ButtonGroup(
  Button("Day", inputId = "range_day"),
  ButtonGroupSection(Text("or", size = "xs")),
  Button("Week", inputId = "range_week")
)
```

## ActionIcon

<https://mantine.dev/core/action-icon/> — an icon-only button, same
click-counter semantics as `Button()`.

```r
ActionIcon(inputId = "like_btn", variant = "light", color = "red", IconHeart(size = 18))
```

`ActionIcon.Group` (`ActionIconGroup()`) and `ActionIconGroupSection()`
work exactly like their `Button.Group` counterparts above.

## CloseButton

<https://mantine.dev/core/close-button/> — a small "x" icon button, used
throughout the package internally (e.g. `Modal()`'s own close control)
and available standalone for your own dismissible UI.

```r
Group(Text("Dismissible notice"), CloseButton(inputId = "dismiss_btn", "aria-label" = "Dismiss"))
```

## CopyButton

<https://mantine.dev/core/copy-button/> — copies `value` to the
clipboard and shows a "Copied" state for a few seconds.

```r
CopyButton(value = "https://mantine.dev", label = "Copy link", copiedLabel = "Copied!")
```

## FileButton

<https://mantine.dev/core/file-button/> — a button that opens the
native file picker on click. Like `FileInput()`, only file *metadata*
(name/size/type) is reported to Shiny, never the file content.

```r
FileButton(inputId = "avatar_upload", label = "Upload avatar", accept = "image/*")
```

## UnstyledButton

<https://mantine.dev/core/unstyled-button/> — a `<button>` with all
default Mantine styling stripped, for building fully custom
clickable elements (e.g. a card that's entirely clickable) while keeping
proper button semantics/accessibility. Purely a passthrough — pair it
with `mantineElement()`-level composition if you need click reporting
beyond what the components above already cover.

```r
UnstyledButton(
  style = list(display = "block", width = "100%", padding = "12px", borderRadius = "8px"),
  Group(Avatar(radius = "xl"), Text("Custom clickable row"))
)
```

## Where to go next

- `vignette("core-navigation")` — `NavLink()`, `Menu()`, `Tabs()`, and
  other click-driven navigation components.
- `inst/examples/buttons-app.R` — every recipe from
  [ui.mantine.dev/category/buttons](https://ui.mantine.dev/category/buttons/)
  reproduced end-to-end.
