---
title: "Core: Layout"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Layout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Layout" category](https://mantine.dev/core/app-shell/)
on mantine.dev/core: structural/positioning components. Every wrapper here
is a thin, direct pass-through to the underlying Mantine component — any
prop documented on the linked mantine.dev page works exactly the same way
from R, just written as a named argument instead of JSX camelCase.

## AppShell

<https://mantine.dev/core/app-shell/> — responsive application shell with
a header, collapsible navbar, contextual aside and footer, all defined
declaratively via layout props rather than manual CSS positioning.

```r
MantineProvider(
  AppShell(
    header = list(height = 60),
    navbar = list(width = 260, breakpoint = "sm", collapsed = list(mobile = TRUE)),
    aside = list(width = 260, breakpoint = "md", collapsed = list(mobile = TRUE)),
    footer = list(height = 50),
    padding = "md",

    AppShellHeader(
      Group(h = "100%", px = "md", navbarBurger("nav_opened", hiddenFrom = "sm"), Title("My app", order = 3))
    ),
    AppShellNavbar(
      p = "md",
      navLinkItem("nav_click", "home", "Home", leftSection = IconHome2(size = 16)),
      navLinkItem("nav_click", "settings", "Settings", leftSection = IconSettings(size = 16))
    ),
    AppShellMain(Text("Page content goes here")),
    AppShellAside(p = "md", Text("Contextual aside")),
    AppShellFooter(p = "md", Text("Footer", size = "xs"))
  )
)
```

`navbarBurger()` (a thin wrapper reporting clicks like `Button()`) is
meant to toggle the navbar's mobile-collapsed state; wire it up with
`updateMantineProps(session, "app_shell_id", navbar = list(collapsed = list(mobile = !opened)))`
if you give the `AppShell()` itself a `mantineId`. See the full working
demo: `shiny::runApp(system.file("examples/appshell-app.R", package = "shiny.mantine"))`.

## AspectRatio

<https://mantine.dev/core/aspect-ratio/> — keeps a fixed width/height
ratio for its single child (e.g. embedding a responsive video or map).

```r
AspectRatio(
  ratio = 16 / 9,
  w = 320,
  tags$iframe(src = "https://www.youtube.com/embed/dQw4w9WgXcQ", frameborder = 0)
)
```

## Center

<https://mantine.dev/core/center/> — centers its content horizontally
and vertically with flexbox; pass `inline = TRUE` to use `inline-flex`
instead of `flex` (useful inside text flow).

```r
Center(h = 200, Stack(align = "center", Loader(), Text("Loading...")))
```

## Container

<https://mantine.dev/core/container/> — centers content horizontally and
constrains its max width, the standard outer wrapper for page content.

```r
Container(size = "sm", py = "xl", Title("Centered narrow content", order = 2))
```

## Flex

<https://mantine.dev/core/flex/> — a flexbox container with shorthand
props for every common flex CSS property (`direction`, `wrap`, `gap`,
`align`, `justify`), for arrangements `Group()`/`Stack()` don't cover
(e.g. `direction` that changes per breakpoint).

```r
Flex(
  direction = list(base = "column", sm = "row"),
  gap = list(base = "sm", sm = "lg"),
  justify = "flex-start",
  align = list(base = "stretch", sm = "center"),
  Button("First", inputId = "flex_btn1"),
  Button("Second", inputId = "flex_btn2"),
  Button("Third", inputId = "flex_btn3")
)
```

## Grid

<https://mantine.dev/core/grid/> — a 12-column responsive grid system
(`GridCol()`'s `span` accepts a number, `"auto"`, `"content"`, or a
per-breakpoint list).

```r
Grid(
  GridCol(span = 8, Paper(withBorder = TRUE, p = "md", Text("Main content (span 8)"))),
  GridCol(span = 4, Paper(withBorder = TRUE, p = "md", Text("Sidebar (span 4)"))),
  GridCol(span = list(base = 12, sm = 6), Paper(withBorder = TRUE, p = "md", Text("Half on tablet+, full on mobile")))
)
```

## Group

<https://mantine.dev/core/group/> — a horizontal flexbox container;
`justify`, `align`, `gap`, `wrap` map directly to their flexbox
equivalents.

```r
Group(justify = "space-between", Text("Left side"), Button("Right side", inputId = "group_btn"))
```

## SimpleGrid

<https://mantine.dev/core/simple-grid/> — a responsive grid with equal
column widths, simpler than `Grid()` when you don't need column
spanning, just "N columns, fewer on small screens":

```r
SimpleGrid(
  cols = list(base = 1, sm = 2, lg = 3),
  spacing = "lg",
  Card(withBorder = TRUE, Text("Card A")),
  Card(withBorder = TRUE, Text("Card B")),
  Card(withBorder = TRUE, Text("Card C"))
)
```

## Space

<https://mantine.dev/core/space/> — an empty horizontal or vertical
spacer, an explicit alternative to margin props.

```r
Group(Text("Left"), Space(w = "xl"), Text("Right, with extra horizontal gap"))
```

## Splitter

<https://mantine.dev/core/splitter/> — resizable panes the user can drag
to resize, composed of `Splitter()` (the root, controls `orientation`)
and `SplitterPanel()` (each pane, with optional `defaultSize`/`minSize`).

```r
Splitter(
  orientation = "horizontal",
  style = list(height = "300px"),
  SplitterPanel(defaultSize = 30, minSize = 20, p = "md", Text("Left pane")),
  SplitterPanel(p = "md", Text("Right pane"))
)
```

## Stack

<https://mantine.dev/core/stack/> — a vertical flexbox container, the
counterpart of `Group()`; almost every example throughout this package's
documentation uses it to arrange components top-to-bottom.

```r
Stack(gap = "md", Title("Section", order = 3), Text("First line"), Button("Action", inputId = "stack_btn"))
```

## Where to go next

- `vignette("core-inputs")` — text/numeric/choice/boolean form inputs.
- `vignette("architecture")` — how these R calls become a React tree.
