---
title: "Core: Overlays"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Overlays}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Overlays" category](https://mantine.dev/core/modal/)
on mantine.dev/core. `Modal()`/`Drawer()`/`Dialog()`/`Popover()` are all
controlled from R the same way: a required `mantineId` identifies them
for [`updateMantineProps(session, mantineId, opened = TRUE/FALSE)`](../reference/updateMantineProps.html),
and an optional `inputId` receives `FALSE` when the user closes them
(click outside, Escape, the built-in close button).

## Modal

<https://mantine.dev/core/modal/> — a centered dialog with a dimmed
backdrop.

```r
Stack(
  Button("Open", inputId = "open_btn"),
  Modal(
    mantineId = "demo_modal", inputId = "modal_state", title = "Demo",
    Text("Are you sure you want to continue?"),
    Button("Confirm", inputId = "confirm_btn")
  )
)

# server:
observeEvent(input$open_btn, {
  updateMantineProps(session, "demo_modal", opened = TRUE)
})
```

More than one `Modal()` open at once needs `ModalStack()` instead of
independent ones (they don't otherwise coordinate z-index/focus
trapping):

```r
ModalStack(
  mantineId = "delete_stack",
  Modal("delete_page", inputId = "delete_page_state", title = "Delete this page?",
    Button("Continue", inputId = "go_to_confirm")
  ),
  Modal("confirm_delete", title = "Are you really sure?",
    Button("Yes, delete", inputId = "confirm_delete_btn", color = "red")
  )
)
```

For full control over the internal layout, use the compound form —
`ModalRoot()` + `ModalOverlay()`/`ModalContent()`/`ModalHeader()`/
`ModalTitle()`/`ModalCloseButton()`/`ModalBody()` — which opens/closes
exactly like `Modal()` (and can join a `ModalStack()` too):

```r
ModalRoot(
  "custom_modal", inputId = "custom_modal_state",
  ModalOverlay(),
  ModalContent(
    ModalHeader(ModalTitle("Custom layout"), ModalCloseButton()),
    ModalBody(Text("Modal content"))
  )
)
```

## Drawer

<https://mantine.dev/core/drawer/> — like `Modal()`, but slides in from
a screen edge (`position = "left"/"right"/"top"/"bottom"`). Same API,
including `DrawerStack()` and the `DrawerRoot()` compound form.

```r
Drawer(mantineId = "settings_drawer", inputId = "drawer_state", title = "Settings", position = "right", Text("Drawer content"))
```

## Dialog

<https://mantine.dev/core/dialog/> — a small, non-modal overlay anchored
to a screen corner (no dimmed backdrop), for lightweight prompts.

```r
Dialog(mantineId = "cookie_dialog", inputId = "cookie_state", opened = TRUE, Text("We use cookies."), Button("Accept", inputId = "accept_btn"))
```

## Popover

<https://mantine.dev/core/popover/> — a floating panel anchored to a
target element; compose with `PopoverTarget()`/`PopoverDropdown()`. Use
this specifically when you need to open/close it from R — for a
self-managed client-side dropdown, `HoverCard()`/`Menu()` below are
simpler.

```r
Popover(
  mantineId = "info_popover", inputId = "popover_state",
  PopoverTarget(Button("Info", inputId = "info_btn")),
  PopoverDropdown(Text("Extra information shown on click."))
)
```

## HoverCard

<https://mantine.dev/core/hover-card/> — a floating panel shown on
hover (not click); fully self-managed client-side, no `mantineId`
needed.

```r
HoverCard(
  HoverCardTarget(Anchor(href = "#", "@username")),
  HoverCardDropdown(Text("Preview shown on hover."))
)
```

## Tooltip

<https://mantine.dev/core/tooltip/> — a small label shown on hover;
`TooltipFloating()` follows the cursor instead of anchoring to the
target, `TooltipGroup()` shares open/close delay across several
tooltips.

```r
Tooltip(label = "Settings", ActionIcon(inputId = "settings_icon_btn", IconSettings(size = 18)))
```

## Menu

<https://mantine.dev/core/menu/> — a dropdown menu opened by clicking
its target; compose with `MenuTarget()`/`MenuDropdown()`/`menuItem()`.
Also supports submenus (`MenuSub()`), checkbox/radio items
(`MenuCheckboxItem()`/`MenuRadioGroup()`), an inline search
(`MenuSearch()`), and a right-click context-menu variant
(`MenuContextMenu()`).

```r
Menu(
  MenuTarget(Button("Actions", inputId = "menu_btn")),
  MenuDropdown(
    menuItem("action", "edit", "Edit", leftSection = IconEdit(size = 16)),
    MenuDivider(),
    menuItem("action", "delete", "Delete", leftSection = IconTrash(size = 16), color = "red")
  )
)
```

## Menubar

<https://mantine.dev/core/menubar/> — a desktop-app-style horizontal bar
of always-visible menus (File, Edit, View, ...), distinct from a single
dropdown `Menu()`.

```r
Menubar(
  MenubarMenu(
    MenubarTarget(Button("File", inputId = "file_menu_btn", variant = "subtle")),
    MenubarDropdown(menuItem("file_action", "new", "New"), menuItem("file_action", "open", "Open"))
  ),
  MenubarMenu(
    MenubarTarget(Button("Edit", inputId = "edit_menu_btn", variant = "subtle")),
    MenubarDropdown(menuItem("edit_action", "undo", "Undo"))
  )
)
```

## Affix

<https://mantine.dev/core/affix/> — fixed-position content (e.g. a
"back to top" button) that stays in place while the page scrolls.

```r
Affix(position = list(bottom = 20, right = 20), Button("Back to top", inputId = "back_to_top_btn"))
```

## LoadingOverlay

<https://mantine.dev/core/loading-overlay/> — a full-cover spinner
overlay for a section of the page while it loads.

```r
Box(
  pos = "relative", h = 200,
  LoadingOverlay(mantineId = "table_loading", visible = TRUE),
  Text("Content being loaded...")
)

# server:
observeEvent(input$data_ready, {
  updateMantineProps(session, "table_loading", visible = FALSE)
})
```

## Overlay

<https://mantine.dev/core/overlay/> — a semi-transparent layer over some
content, the building block `Modal()`/`LoadingOverlay()` use internally
for the dimmed backdrop; available standalone for custom overlays.

```r
Box(pos = "relative", Image(src = "https://placehold.co/400x200"), Overlay(color = "#000", opacity = 0.6, zIndex = 1))
```

## FloatingWindow

<https://mantine.dev/core/floating-window/> — a draggable floating
panel positioned anywhere in the viewport by the user; `initialPosition`
sets where it starts (plain `list(top=, left=)`/`list(bottom=, right=)`
data — the window's *live* position while being dragged isn't
readable/settable from the server, since that needs a client-side ref).

```r
FloatingWindow(
  initialPosition = list(top = 20, right = 20),
  style = list(border = "1px solid #ddd", background = "white", padding = "12px", width = 220),
  Text("Drag me by clicking and holding anywhere on me.")
)
```

## FocusTrap / RemoveScroll

Not their own mantine.dev/core pages, but the two mechanisms
`Modal()`/`Drawer()` use internally — `FocusTrap()` keeps keyboard focus
inside its child while `active`, `RemoveScroll()` locks page scrolling
while `enabled` — exposed for building other fully custom overlays that
need the same behavior. See `vignette("core-misc")`.

## FloatingArrow / FloatingIndicator (not wrapped)

Both need a live floating-UI `context` object and/or a target-element
`ref` to position themselves against another element — neither is a
plain, JSON-serializable R value, so they're not wrapped.

## Where to go next

- `vignette("core-misc")` — `FocusTrap()`, `RemoveScroll()`, and other
  miscellaneous components.
- `vignette("extras")` — group inputs, drag-and-drop reordering, and
  button recipes that don't have their own mantine.dev/core page.
