---
title: "Core: Miscellaneous"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Miscellaneous}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Miscellaneous" category](https://mantine.dev/core/box/)
on mantine.dev/core — components that don't fit neatly into the other
nine categories — plus, at the end, the two core-level provider
components and a summary of what's intentionally not wrapped anywhere in
`shiny.mantine`.

## Box

<https://mantine.dev/core/box/> — the most basic building block: a
`<div>` (or any element via `component`) with access to every Mantine
style prop (`p`, `m`, `bg`, `c`, ...). Most components in this package
are themselves built on top of it.

```r
Box(p = "md", bg = "gray.1", style = list(borderRadius = 8), Text("A styled box"))
```

## Collapse

<https://mantine.dev/core/collapse/> — animated show/hide for its
children; `opened` is controlled entirely from R via `mantineId` +
`updateMantineProps()` (there's no click target built in — pair it with
a `Button()`).

```r
Stack(
  Button("Toggle details", inputId = "toggle_btn"),
  Collapse(mantineId = "details", opened = FALSE, Text("Collapsible content revealed here."))
)

# server:
observeEvent(input$toggle_btn, {
  updateMantineProps(session, "details", opened = input$toggle_btn %% 2 == 1)
})
```

## Divider

<https://mantine.dev/core/divider/> — a horizontal (or vertical) rule,
optionally with a centered `label`.

```r
Stack(Text("Section one"), Divider(label = "OR", labelPosition = "center"), Text("Section two"))
```

## Marquee

<https://mantine.dev/core/marquee/> — a horizontally scrolling ticker of
content, for announcements/logos.

```r
Marquee(Text("Breaking news: shiny.mantine now covers every Mantine core component!"))
```

## Paper

<https://mantine.dev/core/paper/> — a bordered/shadowed surface, the
plain content container `Card()` builds on; used throughout this
package's own examples as the default "box around a demo".

```r
Paper(withBorder = TRUE, shadow = "sm", radius = "md", p = "lg", Text("Content inside a Paper"))
```

## Portal

<https://mantine.dev/core/portal/> — renders its children into a
different part of the DOM (`document.body` by default), the mechanism
`Modal()`/`Menu()`/`Tooltip()` use internally to escape any parent's
`overflow: hidden`/`z-index` stacking context. Rarely needed directly.

```r
Portal(Text("Rendered at the end of <body>, not in its original position in the tree."))
```

## ScrollArea

<https://mantine.dev/core/scroll-area/> — a styled scrollable container
with custom (not browser-native) scrollbars. `ScrollAreaAutosize()`
grows with content up to `mah` before scrolling; `NativeScrollArea()`
opts back into plain OS scrollbars while keeping the same API.

```r
ScrollArea(h = 200, Stack(lapply(1:30, function(i) Text(paste("Line", i)))))
```

## Scroller

<https://mantine.dev/core/scroller/> — a horizontally-scrollable
container with prev/next arrow controls, e.g. for a row of chips or
thumbnails that doesn't fit the available width.

```r
Scroller(
  Group(gap = "xs", wrap = "nowrap", lapply(1:20, function(i) Badge(paste("Item", i))))
)
```

## Transition

<https://mantine.dev/core/transition/> — mounts/unmounts a child with an
animation (fade, slide, scale, ...), controlled by `mounted`.

```r
Stack(
  Button("Toggle", inputId = "transition_btn"),
  Transition(mantineId = "fade_box", mounted = TRUE, transition = "fade", Paper(withBorder = TRUE, p = "md", Text("Fades in/out")))
)

# server:
observeEvent(input$transition_btn, {
  updateMantineProps(session, "fade_box", mounted = input$transition_btn %% 2 == 1)
})
```

## VisuallyHidden

<https://mantine.dev/core/visually-hidden/> — visually hides content
while keeping it available to screen readers (e.g. a label for an
icon-only button that also needs visible sighted-user context
elsewhere).

```r
ActionIcon(inputId = "search_icon_btn", IconSearch(size = 18), VisuallyHidden("Search"))
```

## FocusTrap / RemoveScroll

Not on mantine.dev/core's own "Miscellaneous" page, but closely related
in spirit — the focus-trapping/scroll-locking mechanisms `Modal()`/
`Drawer()` use internally (see `vignette("core-overlays")`), exposed for
building other fully custom overlays:

```r
FocusTrap(
  active = TRUE,
  Stack(
    FocusTrapInitialFocus(), # marks which element should receive focus first
    TextInput(inputId = "first_field", label = "Focused first"),
    TextInput(inputId = "second_field", label = "Second field")
  )
)
RemoveScroll(enabled = TRUE, Paper(withBorder = TRUE, p = "md", Text("Page behind this can't be scrolled while enabled.")))
```

## Provider components

`MantineProvider()` — required, always outermost, see
`vignette("getting-started")` — aside, the two other core-level
providers:

`DirectionProvider()` sets `"ltr"`/`"rtl"` reading direction for
everything nested inside:

```r
MantineProvider(DirectionProvider(dir = "rtl"), Text("Right-to-left content"))
```

`MantineThemeProvider(theme = list(...))` updates the theme *object*
exposed via Mantine's `useMantineTheme()` hook for custom code that
explicitly reads it. **It does not restyle standard components**
(`Button()`, `Badge()`, ...): those get their colors from CSS variables
that only the outermost `MantineProvider()` injects once, so nesting a
`MantineThemeProvider()` with a different `primaryColor` will *not*
change a nested `Button()`'s color:

```r
Group(
  Button("Outside the provider", inputId = "btn_a"),
  MantineThemeProvider(theme = list(primaryColor = "red"), Button("Inside (still the same color)", inputId = "btn_b"))
)
```

If you need a visually different theme for part of a page, that
currently requires a second, independent `MantineProvider()` (with its
own color-scheme-sync caveat — see the README's "Known limitations").

## Intentionally out of scope

A handful of `@mantine/core` exports have no R wrapper, each for a
concrete architectural reason rather than "not gotten to yet" (every
`mantine.dev/core` page was checked against this package's exports):

- **`Combobox` internals** (`ComboboxDropdown`, `ComboboxOption`, ...) —
  driven by Mantine's own `useCombobox()` hook, which can't be expressed
  through R's data-only props; every component built on top of it
  (`Select()`, `MultiSelect()`, ...) is exposed at the higher level — see
  `vignette("core-combobox")`.
- **`FloatingArrow`/`FloatingIndicator`** — need a live floating-UI
  `context` object and/or a target-element `ref`, neither of which is a
  plain, JSON-serializable R value.
- **SSR/Next.js-specific helpers** (`ColorSchemeScript`,
  `HeadlessMantineProvider`, `InlineStyles`) — not applicable to Shiny's
  client-side-only mount architecture.
- **`TableOfContents`/`OverflowList`** — require scanning real page DOM
  headings or render-prop-heavy overflow logic, both a poor fit for a
  declarative R wrapper.
- **`Calendar`** and its sub-parts (`Month`, `Day`, `WeekdaysRow`, ...) —
  the low-level primitive `DatePicker()`/`DatePickerInput()` are built
  on; customization goes through function-based `getDayProps`, a poor
  fit for R's data-only props.

## Where to go next

- `vignette("core-layout")` — where the tour of mantine.dev/core starts.
- `vignette("satellite-packages")` — the separate npm packages
  (`@mantine/dates`, `notifications`, `modals`, `spotlight`, `charts`,
  `code-highlight`, `nprogress`, `tiptap`, `dropzone`, `carousel`).
- `vignette("architecture")` — how every component here is serialized
  and kept in sync with Shiny under the hood.
