---
title: "Core: Navigation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Core: Navigation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

An R rewrite of the ["Navigation" category](https://mantine.dev/core/anchor/)
on mantine.dev/core. `Menu()`/`Menubar()` are covered in
`vignette("core-overlays")` instead, matching mantine.dev's own
categorization of them as overlays, not navigation.

## Anchor

<https://mantine.dev/core/anchor/> — a styled `<a>` link; also usable
`component = "button"`-style for a link-styled clickable action.

```r
Anchor(href = "https://mantine.dev", target = "_blank", "Mantine documentation")
```

## Breadcrumbs

<https://mantine.dev/core/breadcrumbs/> — a navigation trail
(Home > Section > Page), separating its children with `separator`.

```r
Breadcrumbs(Anchor(href = "#", "Home"), Anchor(href = "#", "Settings"), Text("Profile"))
```

## Burger

<https://mantine.dev/core/burger/> — the hamburger-menu icon button that
animates between open/closed states, typically toggling a navbar's
mobile visibility. `navbarBurger()` is a small convenience wrapper that
also reports each click's resulting `opened` value to Shiny (the plain
`Burger()` alone only reports a click counter, like `Button()`):

```r
navbarBurger(inputId = "nav_opened", opened = FALSE, hiddenFrom = "sm")

# server:
observeEvent(input$nav_opened, {
  updateMantineProps(session, "app_shell", navbar = list(collapsed = list(mobile = !input$nav_opened)))
})
```

## NavLink

<https://mantine.dev/core/nav-link/> — a clickable navigation row
(label, optional icon/description, active state), the building block of
sidebar navigation. `navLinkItem()` wraps it with the click-reporting
pattern shared by every menu-like item in this package:

```r
Stack(
  navLinkItem("nav_click", "dashboard", "Dashboard", leftSection = IconLayoutDashboard(size = 16)),
  navLinkItem("nav_click", "settings", "Settings", leftSection = IconSettings(size = 16), active = TRUE)
)
```

Nested inside `Pages()`, `navLinkItem()`'s `pageValue` also drives which
`Page()` is shown client-side with no server round-trip — see
`vignette("getting-started")`'s AppShell example.

## Pagination

<https://mantine.dev/core/pagination/> — page-number controls,
stateful: `input[[inputId]]` is the current page.

```r
Pagination(inputId = "page", total = 10, value = 1, siblings = 1, boundaries = 1)
```

For a fully custom arrangement of the prev/next/first/last controls
instead of the default layout, use the compound form —
`PaginationRoot()` + `PaginationFirst()`/`PaginationPrevious()`/
`PaginationItems()`/`PaginationNext()`/`PaginationLast()`/
`PaginationDots()`/`PaginationControl()`/`PaginationLabel()` — which
read the current page from `PaginationRoot()` automatically:

```r
PaginationRoot(
  inputId = "page2", total = 10,
  Group(PaginationFirst(), PaginationPrevious(), PaginationItems(), PaginationNext(), PaginationLast())
)
```

## Stepper

<https://mantine.dev/core/stepper/> — a multi-step wizard indicator.
Unlike most inputs, the active step is **not** synced automatically on
click — `StepperStep()`'s click is reported as an "event" (like a
button), and moving to the next step is entirely up to your own
server-side logic (e.g. after validating the current step), advanced
with `updateMantineProps()`:

```r
Stepper(
  mantineId = "wizard", inputId = "wizard_click", active = 0,
  StepperStep(label = "Account", description = "Create your account"),
  StepperStep(label = "Details", description = "Add your details"),
  StepperCompleted(Text("Setup complete!"))
)

# server:
observeEvent(input$wizard_click, {
  updateMantineProps(session, "wizard", active = input$wizard_click + 1)
})
```

## Tabs

<https://mantine.dev/core/tabs/> — tabbed navigation; stateful,
`input[[inputId]]` is the active tab's `value`. When nested inside
`Pages()`, it also drives which `Page()` is visible (see
`vignette("getting-started")`).

```r
Tabs(
  inputId = "active_tab", value = "account",
  TabsList(TabsTab("account", "Account"), TabsTab("security", "Security")),
  TabsPanel("account", Text("Account settings")),
  TabsPanel("security", Text("Security settings"))
)
```

## Tree

<https://mantine.dev/core/tree/> — hierarchical, expandable/collapsible
data (e.g. a file explorer). Selecting a node reports its `value` to
Shiny as an event, the same pattern as a menu item click.

```r
Tree(
  inputId = "tree_click",
  data = list(
    list(value = "src", label = "src", children = list(
      list(value = "src/app.R", label = "app.R"),
      list(value = "src/utils.R", label = "utils.R")
    )),
    list(value = "readme", label = "README.md")
  )
)
```

## TableOfContents (not wrapped)

<https://mantine.dev/core/table-of-contents/> requires scanning real
page DOM headings via `getRootElement`/`getControlProps` callbacks — a
poor fit for a declarative, data-only R wrapper. Not covered.

## Where to go next

- `vignette("core-overlays")` — `Menu()`/`Menubar()` and every overlay
  component (modals, drawers, popovers, tooltips, ...).
- `vignette("core-buttons")` — plain clickable buttons and icon buttons.
