---
title: "Extras & Recipes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extras & Recipes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

Components covered by the `core-*` vignettes each have their own
`mantine.dev/core/...` page. This vignette covers what doesn't: group-input
variants that are sub-sections of an existing mantine.dev page rather than
their own, two `shiny.mantine`-only drag-and-drop reordering components, and
a handful of pure-R "recipe" buttons composed on top of `Button()` (the kind
shown on [ui.mantine.dev](https://ui.mantine.dev), not mantine.dev/core).

## Group inputs

<https://mantine.dev/core/checkbox/#checkboxgroup> and
<https://mantine.dev/core/radio/#radiogroup> — `Checkbox()`/`Switch()` are
single booleans; for an array-valued *group* of them, use the dedicated
`CheckboxGroup()`/`SwitchGroup()` plus their own `CheckboxGroupItem()`/
`SwitchGroupItem()` children (never the standalone `Checkbox()`/`Switch()`,
which manage their own independent checked state and would conflict with the
group's):

```r
Stack(
  CheckboxGroup(
    inputId = "fruits", label = "Pick fruits", value = character(0),
    CheckboxGroupItem(value = "apple", label = "Apple"),
    CheckboxGroupItem(value = "banana", label = "Banana")
  ),
  SwitchGroup(
    inputId = "features", label = "Enable features", value = character(0),
    SwitchGroupItem(value = "beta", label = "Beta features"),
    SwitchGroupItem(value = "telemetry", label = "Telemetry")
  )
)

# server:
observe(print(input$fruits))  # e.g. c("apple", "banana")
```

`RadioGroup()`/`ChipGroup()` (covered in `vignette("core-inputs")`) already
follow this same pattern, but nest the plain `Radio()`/`Chip()` directly
since those don't have a separate `checked` state of their own to conflict
with.

<https://mantine.dev/core/checkbox/#checkboxcard> — `CheckboxCard()`/
`RadioCard()` are card-styled selectable items: visually a bordered,
clickable card instead of a small checkbox/radio circle, nested the same way
inside `CheckboxGroup()`/`RadioGroup()`:

```r
RadioGroup(
  inputId = "plan", label = "Choose a plan", value = "free",
  RadioCard(value = "free", Text("Free")),
  RadioCard(value = "pro", Text("Pro"))
)
```

`RadioIndicator()`/`CheckboxIndicator()` are the purely decorative "checked"
look `Radio()`/`Checkbox()` render internally, usable without a real input
behind them — e.g. a read-only status list:

```r
Group(RadioIndicator(checked = TRUE), Text("Step 1 complete"))
```

## Drag-and-drop reordering (`shiny.mantine` extras)

Not mantine.dev components — built on `@hello-pangea/dnd` to fill a gap
Mantine core itself doesn't cover. After each reorder, `input[[inputId]]`
receives the new order as a vector of `value`.

`SortableList()` reorders a plain list of `list(value=, label=)` items,
optionally restricting the drag start to a handle icon:

```r
SortableList(
  "elements", withHandle = TRUE,
  items = list(
    list(value = "H", label = "Hydrogen"),
    list(value = "He", label = "Helium")
  )
)

# server:
observe(print(input$elements))  # e.g. c("He", "H") after dragging Helium to the top
```

`SortableTable()` is the same idea for full table rows (`cells` = the
values for each column):

```r
SortableTable(
  "periodic_rows",
  columns = c("Symbol", "Name"),
  items = list(
    list(value = "H", cells = list("H", "Hydrogen")),
    list(value = "He", cells = list("He", "Helium"))
  )
)
```

## Button recipes

Pure R compositions on top of `Button()`/`ActionIcon()` for patterns shown
on [ui.mantine.dev/category/buttons](https://ui.mantine.dev/category/buttons/)
— not separate mantine.dev/core pages, so not covered in
`vignette("core-buttons")`. See `inst/examples/buttons-app.R` for a runnable
demo of all of them together.

`ColorSchemeToggle()` flips the whole `MantineProvider()`'s light/dark
theme:

```r
ColorSchemeToggle(inputId = "scheme_toggle")
```

`ButtonWithMenu()` opens a dropdown menu on click, items passed as
`menuItem()`:

```r
ButtonWithMenu(
  "Create new",
  menuItem("create_action", "project", "Project"),
  menuItem("create_action", "folder", "Folder")
)
```

`SplitButton()` pairs a primary action with a small arrow opening a menu of
alternatives:

```r
SplitButton(
  "Send", inputId = "send_btn",
  menuItem("send_action", "now", "Send now"),
  menuItem("send_action", "schedule", "Schedule send")
)
```

`LoadingProgressButton()` shows a simulated progress bar over itself on
click; `input[[inputId]]` receives `TRUE` when it reaches 100%:

```r
LoadingProgressButton("Upload", inputId = "upload_done", loadingLabel = "Uploading...")
```

`SocialButton()` is a preconfigured icon + color + label for a common
provider (`"google"`, `"twitter"`, `"facebook"`, `"github"`, `"discord"`) —
copy and adapt its source in `R/Buttons.R` if you need another provider:

```r
Stack(
  SocialButton("google", inputId = "google_login"),
  SocialButton("github", inputId = "github_login")
)
```

## Where to go next

- `vignette("core-layout")` — where the tour of mantine.dev/core starts.
- `vignette("core-inputs")` — `RadioGroup()`/`ChipGroup()`, the group inputs
  that don't need their own `*Item()` wrapper.
- `vignette("architecture")` — how every component here is serialized and
  kept in sync with Shiny under the hood.
