---
title: "Getting Started with shiny.mantine"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with shiny.mantine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

## What is shiny.mantine?

`shiny.mantine` brings [Mantine UI v9](https://mantine.dev) — a modern
React component library covering inputs, layout, navigation, overlays,
charts, and more — into Shiny apps. Every component is a plain R function
that returns a `mantine_element`, composed the same way you'd nest
`htmltools` tags, and every stateful component keeps `input[[inputId]]`
in sync automatically.

This vignette covers installation and your first app. For the concepts
behind the package (how props are serialized, how updates flow from R
back to already-mounted components, the reactive output system), see
`vignette("architecture", package = "shiny.mantine")`. For the ten
Mantine satellite packages (dates, notifications, modals, spotlight,
charts, code-highlight, nprogress, tiptap, dropzone, carousel), see
`vignette("satellite-packages", package = "shiny.mantine")`. For a tour of
every `@mantine/core` component, organized the same way as
[mantine.dev/core](https://mantine.dev/core/), start with
`vignette("core-layout", package = "shiny.mantine")`.

## Installing the package

`shiny.mantine` ships with a *pre-built* JavaScript bundle
(`inst/www/mantine.js`), so installing the R package from a release is
enough to use it — no Node.js toolchain required at install time:

```r
# from a local checkout or a released source tarball
devtools::install()
```

### Building the JS bundle from source

If you're working from a development checkout where the bundle needs to
be (re)built — e.g. after pulling changes to `js/src/index.js`, or after
editing the component generator's manifest — you'll need Node.js and npm:

```
cd js
npm install
npm run build
```

This compiles `js/src/index.js` (which imports the whole Mantine core
library plus all ten satellite packages, its own bundled copy of React
19, and a handful of smaller integrations) into `inst/www/mantine.js` via
webpack. The bundle is intentionally self-contained — see
`vignette("architecture")` for why it bundles its own React copy instead
of sharing one with `shiny.react`.

After rebuilding the JS bundle, regenerate the R documentation and
reinstall:

```r
devtools::document()
devtools::install()
```

## Your first app

Every `shiny.mantine` UI needs exactly one `MantineProvider()` wrapping
everything — it supplies the theme and color-scheme context that every
other component depends on:

```r
library(shiny)
library(shiny.mantine)

ui <- fluidPage(
  MantineProvider(
    Stack(
      Title("Hello, Mantine"),
      TextInput(inputId = "name", label = "Your name", placeholder = "Ada Lovelace"),
      Button("Greet me", inputId = "greet_btn"),
      Text(textOutput("greeting"))
    )
  )
)

server <- function(input, output, session) {
  output$greeting <- renderText({
    req(input$greet_btn > 0)
    paste("Hello,", input$name %||% "stranger", "!")
  })
}

shinyApp(ui, server)
```

A few things to notice:

- **`TextInput(inputId = "name", ...)`** behaves exactly like
  `shiny::textInput()`: `input$name` updates on every keystroke.
- **`Button(..., inputId = "greet_btn")`** behaves like
  `shiny::actionButton()`: `input$greet_btn` is a click counter.
- Mixing a native Shiny output (`textOutput()`) *inside* a
  `shiny.mantine` component (`Text()`) works fine — it's the reverse
  (nesting a native Shiny binding as a **direct structural child of**
  `MantineProvider()`'s own markup) that doesn't, since that content is
  injected via `dangerouslySetInnerHTML` and Shiny won't re-scan it for
  bindings. Keep native Shiny inputs/outputs either fully outside
  `MantineProvider()`, or nested inside a `shiny.mantine` component like
  `Text()`/`Card()` (as shown above), not as siblings passed directly to
  `MantineProvider()`'s own `...`.

Everything should render at its normal, intended size out of the box,
regardless of whether your UI uses `fluidPage()` (which loads Shiny's
bundled Bootstrap 3 CSS, setting the page's root font-size to 10px
instead of the browser's 16px default) or a plain `tagList()` (which
doesn't). `MantineProvider()` measures the real root font-size at mount
time and compensates automatically either way — see "Font-size scaling
under Shiny's Bootstrap CSS" in the README — but it's worth knowing about
if you ever pass your own `theme` with an explicit `scale`, or set
`fixShinyFontScale = FALSE`.

## A fuller example: AppShell with page navigation

The package ships a demo with a complete responsive layout — header,
collapsible navbar, contextual aside, footer — and a client-side page
router so navigating between "pages" needs no server round-trip:

```r
shiny::runApp(system.file("examples/appshell-app.R", package = "shiny.mantine"))
```

Every [ui.mantine.dev](https://ui.mantine.dev/) category also has its own
runnable demo in `inst/examples/` — see the README's "Examples per
category" table, or list them with:

```r
list.files(system.file("examples", package = "shiny.mantine"))
```

## Where to go next

- `vignette("architecture")` — how the R-to-React serialization works,
  the generic `updateMantineProps()` channel, and `mantineOutput()`/`renderMantine()`.
- `vignette("core-layout")` — the first of ten `core-*` vignettes
  touring every `@mantine/core` component, one per mantine.dev/core
  category.
- `vignette("satellite-packages")` — dates, notifications, modals,
  spotlight, charts, code-highlight, nprogress, the rich text editor,
  dropzone, and carousel.
- `vignette("extras")` — group inputs, drag-and-drop reordering, and
  button recipes that don't have their own mantine.dev/core page.
