---
title: "Satellite Packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Satellite Packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

Beyond Mantine's core component library (see the `core-*` vignettes,
starting with `vignette("core-layout")`), `shiny.mantine` bundles every
separate Mantine v9 npm package: dates, notifications, modals,
spotlight, charts, code-highlight, nprogress, a rich text editor
(tiptap), drag-and-drop file uploads (dropzone), and an image/content
carousel (carousel). This vignette walks through each with a runnable
example.

## `@mantine/dates`

Dates travel between R and JS as plain `"YYYY-MM-DD"` strings — Mantine
v9 itself represents dates this way (`DateStringValue = string`, not a JS
`Date` object), which makes them trivially JSON-serializable. Every date
input accepts either an R `Date`/`POSIXct` (converted automatically via
the internal `toDateString()` helper) or an already-formatted string:

```r
ui <- fluidPage(
  MantineProvider(
    Stack(
      DateInput("birth_date", label = "Date of birth", value = Sys.Date()),
      DatePickerInput("appointment", label = "Appointment"),
      TimeInput("meeting_time", label = "Meeting time"),
      DateTimePicker("event_at", label = "Event date & time")
    )
  )
)

server <- function(input, output, session) {
  observe(print(input$birth_date))  # a plain "YYYY-MM-DD" string
}

shinyApp(ui, server)
```

`DatesProvider()` sets locale/`firstDayOfWeek` for every date component
nested inside it. Also available: `MonthPickerInput()`/`YearPickerInput()`
(popover month/year-only selection) and their always-visible inline
counterparts `MonthPicker()`/`YearPicker()` (like `DatePicker()`, but for
month/year granularity), `TimePicker()`/`TimeGrid()` (dropdown/grid time
pickers, distinct from the native `TimeInput()`), `MiniCalendar()`
(compact single-week strip), `InlineDateTimePicker()` (like
`DateTimePicker()`, but always visible instead of behind a popover), and
`TimeValue()` (a display-only component that formats a time string/`Date`
for 12h/24h display, e.g. inside a `Text()`).

## `@mantine/notifications`

`Notifications()` (Mantine-styled toasts) needs to be mounted **once** in
the page — typically as a sibling passed to `MantineProvider()` alongside
your main content — then driven imperatively from the server:

```r
ui <- fluidPage(
  MantineProvider(
    Notifications(position = "top-right"),
    Button("Save", inputId = "save_btn")
  )
)

server <- function(input, output, session) {
  observeEvent(input$save_btn, {
    showMantineNotification(session, title = "Saved", message = "Your changes were saved.", color = "green")
  })
}

shinyApp(ui, server)
```

Use `hideMantineNotification(session, id)` to dismiss a specific
notification early (pass a matching `id = "..."` when showing it).

## `@mantine/modals`

`ModalsProvider()` enables an imperative confirm/generic modal API, useful
when you don't want to pre-declare a `Modal()` in the UI:

```r
ui <- fluidPage(
  MantineProvider(
    ModalsProvider(
      Button("Delete item", inputId = "delete_btn", color = "red")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$delete_btn, {
    openMantineConfirmModal(
      session, inputId = "confirm_delete",
      title = "Confirm deletion",
      children = "This action cannot be undone.",
      labels = list(confirm = "Delete", cancel = "Cancel"),
      confirmProps = list(color = "red")
    )
  })
  observeEvent(input$confirm_delete, {
    if (isTRUE(input$confirm_delete)) {
      # ... actually delete the item ...
    }
  })
}

shinyApp(ui, server)
```

**Important**: if you use both `Notifications()`/`ModalsProvider()` and
your own overlays, `MantineProvider()` must always be the *outermost*
call — never nest it inside `ModalsProvider()`.

## `@mantine/spotlight`

A command palette (opened with Ctrl/Cmd+K by default). Each action
reports its own `id` to `input[[inputId]]` when selected — the same
pattern as `menuItem()`/`navLinkItem()`:

```r
ui <- fluidPage(
  MantineProvider(
    Spotlight(
      inputId = "spotlight_action",
      actions = list(
        list(id = "home", label = "Go to Home", description = "Main page"),
        list(id = "settings", label = "Go to Settings")
      )
    ),
    Text("Press Ctrl/Cmd+K to open the command palette")
  )
)

server <- function(input, output, session) {
  observeEvent(input$spotlight_action, {
    message("Navigating to: ", input$spotlight_action)
  })
}

shinyApp(ui, server)
```

For grouped actions (or a fully custom search/empty/footer layout) that
the flat `actions` list above doesn't allow, use `SpotlightRoot()` +
`SpotlightSearch()`/`SpotlightActionsList()`/`SpotlightActionsGroup()`/
`SpotlightAction()`/`SpotlightEmpty()`/`SpotlightFooter()` — the same
global command-palette instance, no store management needed from R.
Unlike `Spotlight()`, the compound form doesn't filter actions by the
search query itself: `SpotlightSearch()` only renders the search box,
matching typed text against actions (and conditionally including
`SpotlightEmpty()` only when nothing matches) is up to you:

```r
SpotlightRoot(
  SpotlightSearch(placeholder = "Search..."),
  SpotlightActionsList(
    SpotlightActionsGroup(
      label = "Navigation",
      SpotlightAction("spotlight_choice", "home", label = "Home"),
      SpotlightAction("spotlight_choice", "settings", label = "Settings")
    )
  )
)
```

## `@mantine/charts`

Chart components accept `data` as either a `data.frame` (one row per
point/category — converted automatically into a list of rows) or an
already-prepared list:

```r
sales <- data.frame(month = c("Jan", "Feb", "Mar", "Apr"), revenue = c(120, 150, 90, 200))

ui <- fluidPage(
  MantineProvider(
    Stack(
      LineChart(
        data = sales, dataKey = "month",
        series = list(list(name = "revenue", color = "blue.6"))
      ),
      PieChart(data = data.frame(
        name = c("Desktop", "Mobile", "Tablet"),
        value = c(55, 35, 10),
        color = c("blue.6", "orange.6", "grape.6")
      ))
    )
  )
)

shinyApp(ui, function(input, output, session) {})
```

`BarChart()`/`AreaChart()`/`RadarChart()`/`CompositeChart()` share
`LineChart()`'s `data`/`series` shape; `DonutChart()`/`FunnelChart()` share
`PieChart()`'s `name`/`value`/`color` shape. `RadialBarChart()`/
`BubbleChart()` take `data`/`dataKey` without `series`. `Sparkline()` takes
a plain numeric vector instead of a `data.frame`. `BarsList()` takes a
`data.frame`/list of rows with `name`/`value` columns, rendering each row
as a horizontal bar with its value (no axes) — good for compact
top-N/leaderboard displays. `SunburstChart()` (added in Mantine 9.5) takes
a nested list like `Treemap()` (`list(name=, value=, color=,
children=...)`) and plots it as concentric rings. `BulletChart()` (also
added in 9.5) is a compact KPI chart: an actual `value` plotted against an
optional `target` and a `ranges` data.frame/list of rows (`value` = each
range's upper bound, `color`, optional `label`) shown as background bands.
`ScatterChart()`, `Treemap()`, `Heatmap()`, and `SankeyChart()` have their
own nested `data` shapes (documented in each function's `?help` page)
rather than one flat table, since their charts aren't naturally
one-row-per-point.

```r
BulletChart(
  value = 260000, target = 275000, label = "Revenue",
  ranges = data.frame(
    value = c(150000, 225000, 300000),
    color = c("red.8", "yellow.8", "teal.8"),
    label = c("Poor", "Average", "Good")
  )
)
```

Charts with an unbounded container (e.g. a plain `Stack()` with no fixed
height, as above) need an explicit `h` (height) — Recharts, which every
Mantine chart is built on, silently renders nothing if it can't resolve a
definite pixel height for itself. `LineChart()`/`BarChart()`/`PieChart()`/
etc. happen to ship a sensible default; several of the newer chart types
(`RadarChart()`, `RadialBarChart()`, `BubbleChart()`, `CompositeChart()`,
`ScatterChart()`, `SunburstChart()`) do not, so pass `h = 250` (or
similar) explicitly if one of these renders blank.

Since Mantine 9.5, `LineChart()`/`BarChart()`/`AreaChart()`/
`CompositeChart()`/`ScatterChart()`/`BubbleChart()`/`PieChart()`/
`DonutChart()`/`RadarChart()`/`RadialBarChart()`/`FunnelChart()` support
`accessibilityLayer` (keyboard navigation, `TRUE` by default) and
`LineChart()`/`BarChart()`/`AreaChart()`/`CompositeChart()` support
`withBrush = TRUE` to show a range-selector brush under the chart — both
plain boolean props, passed through like any other.

## `@mantine/code-highlight`

`CodeHighlight()`/`InlineCodeHighlight()` display code with a monospace
font, line numbers, and a copy button, but **without syntax coloring** —
they always use Mantine's `plainTextAdapter` rather than pulling in
`highlight.js`/`shiki`, to keep the bundle from growing further.
`CodeHighlightTabs(code = list(list(fileName = ..., language = ..., code =
...), ...))` displays several snippets as a tabbed file browser, same
adapter. If you
need real syntax highlighting, you'd need to add such an adapter to
`js/src/index.js` yourself.

```r
CodeHighlight(
  code = "server <- function(input, output, session) {\n  # ...\n}",
  language = "r"
)
```

## `@mantine/nprogress`

A slim progress bar pinned to the top of the page — the kind used for
page-navigation loading indicators. Mount `NavigationProgress()` once,
then drive it imperatively:

```r
ui <- fluidPage(
  MantineProvider(
    NavigationProgress(),
    Button("Run task", inputId = "go_btn")
  )
)

server <- function(input, output, session) {
  observeEvent(input$go_btn, {
    startMantineProgress(session)
    setMantineProgress(30, session)
    Sys.sleep(0.5)  # stand-in for real work
    setMantineProgress(70, session)
    Sys.sleep(0.5)
    completeMantineProgress(session)
  })
}

shinyApp(ui, server)
```

`resetMantineProgress()`/`incrementMantineProgress()`/`decrementMantineProgress()`
round out the imperative API.

## `@mantine/tiptap`

`RichTextEditor()` is a rich text editor with **reduced scope** compared
to a full Tiptap setup: bold/italic/underline/strikethrough/highlight/
inline code, H1-H6 headings, blockquotes, lists (incl. task lists),
links, alignment, sub/superscript, and undo/redo — no tables, images,
text color, or collaborative editing.

```r
ui <- fluidPage(
  MantineProvider(
    RichTextEditor("bio", content = "<p>Write something...</p>", placeholder = "Your bio"),
    Button("Reset", inputId = "reset_btn")
  ),
  verbatimTextOutput("html")
)

server <- function(input, output, session) {
  observeEvent(input$reset_btn, {
    updateMantineRichTextEditor(session, "bio", content = "<p></p>")
  })
  output$html <- renderPrint(input$bio)
}

shinyApp(ui, server)
```

`input[[inputId]]` receives the editor's content as HTML on every edit —
treat it as untrusted user input if you ever render it back with
`HTML()`/`dangerouslySetInnerHTML`-style trust, the same way you would any
other free-text field.

Use `controls` to trim down or reorder the toolbar (a list of character
vectors, one per visually separated button cluster):

```r
RichTextEditor(
  "notes", content = "<p></p>",
  controls = list(c("bold", "italic"), c("bulletList", "orderedList"))
)
```

## `@mantine/dropzone`

`Dropzone()` is a drag-and-drop (or click-to-browse) file upload area.
Like `FileInput()`/`FileButton()` in `vignette("core-inputs")`/
`vignette("core-buttons")`, only file *metadata* (name/size/type) is
reported to Shiny, never the file content — pair it with a real
`shiny::fileInput()` alongside it if you need the actual upload.
`DropzoneAccept()`/`DropzoneReject()`/`DropzoneIdle()` render different
content depending on whether the dragged file(s) currently over the
dropzone would be accepted.

```r
ui <- fluidPage(
  MantineProvider(
    Dropzone(
      inputId = "upload", accept = "image/png,image/jpeg", maxSize = 5 * 1024^2,
      DropzoneAccept(Text("Drop the file here")),
      DropzoneReject(Text("File type not accepted", c = "red")),
      DropzoneIdle(Text("Drag an image here, or click to select one"))
    )
  ),
  verbatimTextOutput("files")
)

server <- function(input, output, session) {
  output$files <- renderPrint(input$upload)
}

shinyApp(ui, server)
```

`DropzoneFullScreen()` is the same idea, but instead of a fixed drop
area it listens for drops anywhere on the browser window while its
`active` prop is `TRUE` — typically toggled by a button via
`updateMantineProps(session, mantineId, active = TRUE/FALSE)`:

```r
ui <- fluidPage(
  MantineProvider(
    Button("Activate full-screen drop", inputId = "activate_btn"),
    DropzoneFullScreen(
      inputId = "upload", mantineId = "full_dz", active = FALSE,
      Text("Drop files anywhere on the page")
    )
  )
)

server <- function(input, output, session) {
  active <- reactiveVal(FALSE)
  observeEvent(input$activate_btn, {
    active(!active())
    updateMantineProps(session, "full_dz", active = active())
  })
}

shinyApp(ui, server)
```

## `@mantine/carousel`

`Carousel()`/`CarouselSlide()` — a swipeable/scrollable carousel for
images or any other content, built on Embla Carousel.

```r
ui <- fluidPage(
  MantineProvider(
    Carousel(
      height = 200, slideSize = "33.333%", slideGap = "md", withIndicators = TRUE,
      CarouselSlide(Image(src = "https://placehold.co/400x200?text=1")),
      CarouselSlide(Image(src = "https://placehold.co/400x200?text=2")),
      CarouselSlide(Image(src = "https://placehold.co/400x200?text=3"))
    )
  )
)

server <- function(input, output, session) {}

shinyApp(ui, server)
```

## `@mantine/schedule`

Calendar/scheduling views: `DayView()`, `WeekView()`, `MonthView()`,
`YearView()`, `AgendaView()` (a scrollable event list over a fixed date
range), `MobileMonthView()` (touch-optimized), the resource-grouped
`ResourcesDayView()`/`ResourcesWeekView()`/`ResourcesMonthView()` (events
grouped into rows/columns by e.g. meeting room or staff member), and
`Schedule()` (all of the above behind one component with its own
day/week/month/year switcher). Events are a `data.frame`/list of rows with
`id`, `title`, `start`, `end`, and optionally `color`/`resourceId`
columns — `start`/`end` accept `"YYYY-MM-DD HH:mm:ss"` strings or R
`Date`/`POSIXct` values, converted automatically.

```r
events <- data.frame(
  id = 1:2,
  title = c("Team standup", "Client call"),
  start = c("2026-07-27 09:00:00", "2026-07-27 11:00:00"),
  end = c("2026-07-27 09:30:00", "2026-07-27 12:00:00"),
  color = c("blue", "green")
)

ui <- fluidPage(
  MantineProvider(
    DayView(inputId = "day1", date = Sys.Date(), events = events, startTime = "08:00:00")
  ),
  verbatimTextOutput("clicked")
)

server <- function(input, output, session) {
  output$clicked <- renderPrint(input$day1_event_click)
}

shinyApp(ui, server)
```

Every view reports its currently displayed `date` as `input[[inputId]]`
(the primary value, like every other stateful input in this package); a
handful of other interactions are reported under suffixed input ids —
`input[[paste0(inputId, "_event_click")]]` (clicked event's `id`),
`input[[paste0(inputId, "_slot_click")]]`, `input[[paste0(inputId,
"_slot_select")]]` (a drag-to-select range, needs `withDragSlotSelect =
TRUE`), `input[[paste0(inputId, "_event_drop")]]`/`input[[paste0(inputId,
"_event_resize")]]` (needs `withEventsDragAndDrop`/`withEventResize =
TRUE`) — see `?ScheduleViews` for the full list. Update `date`/`events`
(or any other prop) from the server with `updateMantineProps()` (needs a
`mantineId`).

`ResourcesDayView()`/`ResourcesWeekView()`/`ResourcesMonthView()` also
take a `resources` data.frame/list (`id`/`label` columns), and each
event's `resourceId` should match one of them. Their `intervalMinutes`
prop (default `60`) accepts values dividing evenly into an hour (e.g.
`15`, `30`) *or* whole numbers of hours (e.g. `120`, `240`) — values
above `60` widen each column to span several hours instead of
subdividing within an hour, useful to fit a whole day in view compactly:

```r
ResourcesDayView(
  inputId = "rooms1",
  date = Sys.Date(),
  resources = data.frame(id = c("tokyo", "paris"), label = c("Tokyo", "Paris")),
  events = data.frame(
    id = 1, title = "Team standup",
    start = "2026-07-27 09:00:00", end = "2026-07-27 10:00:00",
    resourceId = "tokyo"
  ),
  intervalMinutes = 240 # 4-hour-wide columns spanning midnight to midnight
)
```

Render-prop customization (`renderEvent`, `renderEventBody`,
`renderResourceLabel`, custom per-event drag/resize predicates, ...) is
JS-callback-based and not exposed — every event is draggable/resizable
whenever the matching `with*` flag is set, and default rendering is used
throughout (the same trade-off already made for `Combobox()`/
`TableOfContents()`/`OverflowList()`; see `vignette("core-misc")`).
