---
title: "Satellite Packages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Satellite Packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shiny)
library(shiny.mantine)
```

Beyond Mantine's core component library, `shiny.mantine` bundles all
eight Mantine v9 satellite packages: dates, notifications, modals,
spotlight, charts, code-highlight, nprogress, and a rich text editor
(tiptap). This vignette walks through each with a runnable example.

## `@mantine/dates`

Dates travel between R and JS as plain `"YYYY-MM-DD"` strings — Mantine
v9 itself represents dates this way (`DateStringValue = string`, not a JS
`Date` object), which makes them trivially JSON-serializable. Every date
input accepts either an R `Date`/`POSIXct` (converted automatically via
the internal `toDateString()` helper) or an already-formatted string:

```r
ui <- fluidPage(
  MantineProvider(
    Stack(
      DateInput("birth_date", label = "Date of birth", value = Sys.Date()),
      DatePickerInput("appointment", label = "Appointment"),
      TimeInput("meeting_time", label = "Meeting time"),
      DateTimePicker("event_at", label = "Event date & time")
    )
  )
)

server <- function(input, output, session) {
  observe(print(input$birth_date))  # a plain "YYYY-MM-DD" string
}

shinyApp(ui, server)
```

`DatesProvider()` sets locale/`firstDayOfWeek` for every date component
nested inside it. Also available: `MonthPickerInput()`/`YearPickerInput()`
(month/year-only selection), `TimePicker()`/`TimeGrid()` (dropdown/grid
time pickers, distinct from the native `TimeInput()`), `MiniCalendar()`
(compact single-week strip), and `InlineDateTimePicker()` (like
`DateTimePicker()`, but always visible instead of behind a popover).

## `@mantine/notifications`

`Notifications()` (Mantine-styled toasts) needs to be mounted **once** in
the page — typically as a sibling passed to `MantineProvider()` alongside
your main content — then driven imperatively from the server:

```r
ui <- fluidPage(
  MantineProvider(
    Notifications(position = "top-right"),
    Button("Save", inputId = "save_btn")
  )
)

server <- function(input, output, session) {
  observeEvent(input$save_btn, {
    showMantineNotification(session, title = "Saved", message = "Your changes were saved.", color = "green")
  })
}

shinyApp(ui, server)
```

Use `hideMantineNotification(session, id)` to dismiss a specific
notification early (pass a matching `id = "..."` when showing it).

## `@mantine/modals`

`ModalsProvider()` enables an imperative confirm/generic modal API, useful
when you don't want to pre-declare a `Modal()` in the UI:

```r
ui <- fluidPage(
  MantineProvider(
    ModalsProvider(
      Button("Delete item", inputId = "delete_btn", color = "red")
    )
  )
)

server <- function(input, output, session) {
  observeEvent(input$delete_btn, {
    openMantineConfirmModal(
      session, inputId = "confirm_delete",
      title = "Confirm deletion",
      children = "This action cannot be undone.",
      labels = list(confirm = "Delete", cancel = "Cancel"),
      confirmProps = list(color = "red")
    )
  })
  observeEvent(input$confirm_delete, {
    if (isTRUE(input$confirm_delete)) {
      # ... actually delete the item ...
    }
  })
}

shinyApp(ui, server)
```

**Important**: if you use both `Notifications()`/`ModalsProvider()` and
your own overlays, `MantineProvider()` must always be the *outermost*
call — never nest it inside `ModalsProvider()`.

## `@mantine/spotlight`

A command palette (opened with Ctrl/Cmd+K by default). Each action
reports its own `id` to `input[[inputId]]` when selected — the same
pattern as `menuItem()`/`navLinkItem()`:

```r
ui <- fluidPage(
  MantineProvider(
    Spotlight(
      inputId = "spotlight_action",
      actions = list(
        list(id = "home", label = "Go to Home", description = "Main page"),
        list(id = "settings", label = "Go to Settings")
      )
    ),
    Text("Press Ctrl/Cmd+K to open the command palette")
  )
)

server <- function(input, output, session) {
  observeEvent(input$spotlight_action, {
    message("Navigating to: ", input$spotlight_action)
  })
}

shinyApp(ui, server)
```

For grouped actions (or a fully custom search/empty/footer layout) that
the flat `actions` list above doesn't allow, use `SpotlightRoot()` +
`SpotlightSearch()`/`SpotlightActionsList()`/`SpotlightActionsGroup()`/
`SpotlightAction()`/`SpotlightEmpty()`/`SpotlightFooter()` — the same
global command-palette instance, no store management needed from R.
Unlike `Spotlight()`, the compound form doesn't filter actions by the
search query itself: `SpotlightSearch()` only renders the search box,
matching typed text against actions (and conditionally including
`SpotlightEmpty()` only when nothing matches) is up to you:

```r
SpotlightRoot(
  SpotlightSearch(placeholder = "Search..."),
  SpotlightActionsList(
    SpotlightActionsGroup(
      label = "Navigation",
      SpotlightAction("spotlight_choice", "home", label = "Home"),
      SpotlightAction("spotlight_choice", "settings", label = "Settings")
    )
  )
)
```

## `@mantine/charts`

Chart components accept `data` as either a `data.frame` (one row per
point/category — converted automatically into a list of rows) or an
already-prepared list:

```r
sales <- data.frame(month = c("Jan", "Feb", "Mar", "Apr"), revenue = c(120, 150, 90, 200))

ui <- fluidPage(
  MantineProvider(
    Stack(
      LineChart(
        data = sales, dataKey = "month",
        series = list(list(name = "revenue", color = "blue.6"))
      ),
      PieChart(data = data.frame(
        name = c("Desktop", "Mobile", "Tablet"),
        value = c(55, 35, 10),
        color = c("blue.6", "orange.6", "grape.6")
      ))
    )
  )
)

shinyApp(ui, function(input, output, session) {})
```

`BarChart()`/`AreaChart()`/`RadarChart()`/`CompositeChart()` share
`LineChart()`'s `data`/`series` shape; `DonutChart()`/`FunnelChart()` share
`PieChart()`'s `name`/`value`/`color` shape. `RadialBarChart()`/
`BubbleChart()` take `data`/`dataKey` without `series`. `Sparkline()` takes
a plain numeric vector instead of a `data.frame`. `ScatterChart()`,
`Treemap()`, `Heatmap()`, and `SankeyChart()` have their own nested `data`
shapes (documented in each function's `?help` page) rather than one flat
table, since their charts aren't naturally one-row-per-point.

Charts with an unbounded container (e.g. a plain `Stack()` with no fixed
height, as above) need an explicit `h` (height) — Recharts, which every
Mantine chart is built on, silently renders nothing if it can't resolve a
definite pixel height for itself. `LineChart()`/`BarChart()`/`PieChart()`/
etc. happen to ship a sensible default; several of the newer chart types
(`RadarChart()`, `RadialBarChart()`, `BubbleChart()`, `CompositeChart()`,
`ScatterChart()`) do not, so pass `h = 250` (or similar) explicitly if one
of these renders blank.

## `@mantine/code-highlight`

`CodeHighlight()`/`InlineCodeHighlight()` display code with a monospace
font, line numbers, and a copy button, but **without syntax coloring** —
they always use Mantine's `plainTextAdapter` rather than pulling in
`highlight.js`/`shiki`, to keep the bundle from growing further.
`CodeHighlightTabs(code = list(list(fileName = ..., language = ..., code =
...), ...))` displays several snippets as a tabbed file browser, same
adapter. If you
need real syntax highlighting, you'd need to add such an adapter to
`js/src/index.js` yourself.

```r
CodeHighlight(
  code = "server <- function(input, output, session) {\n  # ...\n}",
  language = "r"
)
```

## `@mantine/nprogress`

A slim progress bar pinned to the top of the page — the kind used for
page-navigation loading indicators. Mount `NavigationProgress()` once,
then drive it imperatively:

```r
ui <- fluidPage(
  MantineProvider(
    NavigationProgress(),
    Button("Run task", inputId = "go_btn")
  )
)

server <- function(input, output, session) {
  observeEvent(input$go_btn, {
    startMantineProgress(session)
    setMantineProgress(30, session)
    Sys.sleep(0.5)  # stand-in for real work
    setMantineProgress(70, session)
    Sys.sleep(0.5)
    completeMantineProgress(session)
  })
}

shinyApp(ui, server)
```

`resetMantineProgress()`/`incrementMantineProgress()`/`decrementMantineProgress()`
round out the imperative API.

## `@mantine/tiptap`

`RichTextEditor()` is a rich text editor with **reduced scope** compared
to a full Tiptap setup: bold/italic/underline/strikethrough, H1-H4
headings, blockquotes, lists, links, alignment, and undo/redo — no
tables, images, or collaborative editing.

```r
ui <- fluidPage(
  MantineProvider(
    RichTextEditor("bio", content = "<p>Write something...</p>", placeholder = "Your bio"),
    Button("Reset", inputId = "reset_btn")
  ),
  verbatimTextOutput("html")
)

server <- function(input, output, session) {
  observeEvent(input$reset_btn, {
    updateMantineRichTextEditor(session, "bio", content = "<p></p>")
  })
  output$html <- renderPrint(input$bio)
}

shinyApp(ui, server)
```

`input[[inputId]]` receives the editor's content as HTML on every edit —
treat it as untrusted user input if you ever render it back with
`HTML()`/`dangerouslySetInnerHTML`-style trust, the same way you would any
other free-text field.
